---
title: "NP AMR analyses"
author: "Nicholas Croucher"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(magrittr)
require(tidyverse)

```

## Load data from project

Import sample data from project

```{r Process sample data}

parse_sample_file <- function(fn) {
  
  read.table(file = fn,
             sep = "\t",
             comment.char = "",
             header = TRUE)
  
}

process_annotation_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "annotation") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",path)) %>% 
    dplyr::mutate(annotation_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_',npg_run_id)) %>% 
    dplyr::mutate(npg_run_id = gsub(".gff","",npg_run_id)) %>% 
    dplyr::select(-path)
  
}

process_assembly_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "assembly") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/velvet_assembly/contigs.fa","",gsub("/spades_assembly/contigs.fa","",path)))) %>% 
    dplyr::mutate(assembly_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_contigs.fa')) %>% 
    dplyr::select(npg_run_id,assembly_file)
  
}

process_qc_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "qc") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/kraken.report","",path))) %>% 
    dplyr::mutate(qc_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_kraken.report')) %>% 
    dplyr::select(npg_run_id,qc_file)
  
}

# Parse sample files
info_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_info.tsv") %>% 
      dplyr::filter(!grepl("HiC_RMV",sample_supplier_name)) %>%
      dplyr::mutate(project = "6588"),
    parse_sample_file("./sample_data/6310_info.tsv") %>%
      dplyr::mutate(project = "6310")
  )

run_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_runs.tsv") %>% 
      dplyr::select(-name),
    parse_sample_file("./sample_data/6310_runs.tsv") %>% 
      dplyr::select(-name)
  )

fn_6588 <- "./sample_data/6588_files.tsv"
fn_6310 <- "./sample_data/6310_files.tsv"
files_df <-
  dplyr::bind_rows(
    parse_sample_file(fn_6588) %>% process_annotation_file_names(),
    parse_sample_file(fn_6310) %>% process_annotation_file_names()
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_assembly_file_names(),
      parse_sample_file(fn_6310) %>% process_assembly_file_names()
    ),
    by = c("npg_run_id")
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_qc_file_names(),
      parse_sample_file(fn_6310) %>% process_qc_file_names()
    ),
    by = c("npg_run_id")
  )

epi_df <-
  dplyr::bind_rows(
    read.csv("./sample_data/Asian_isolate_data.csv") %>% 
      dplyr::mutate(Description = gsub(" from nasopharyngeal swab","",Description)) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Species = Description) %>% 
      dplyr::mutate(Description = dplyr::if_else(Species == "Mixed bacteria","Mixed bacteria","Single isolate")),
    read.csv("./sample_data/Hampshire_isolate_data.csv") %>% 
      dplyr::mutate(Cohort = Country) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Description = "Single isolate"),
    read.csv("./sample_data/Global-Hinf-metadata-microreact.csv") %>% 
      dplyr::filter(disease == "") %>% 
      dplyr::select(id,country,collection_date) %>% 
      dplyr::filter(country == "Maela") %>% 
      dplyr::rename(Isolate = id,
                    Region = country,
                    Year = collection_date) %>% 
      dplyr::mutate(Sample = Isolate,
                    Region = "Maela",
                    Cohort = "Maela",
                    Country = "Thailand",
                    Species = "Haemophilus influenzae",
                    project = "6310",
                    Year = as.numeric(Year)) %>% 
      dplyr::mutate(npg_run_id = gsub("(_.*?)_", "\\1#", Sample)) %>% 
      dplyr::mutate(Description = "Single isolate")
  ) %>% 
  dplyr::mutate(join_col = dplyr::if_else(project == "6588",Sample,npg_run_id)) %>% 
  dplyr::select(-c(project,npg_run_id,Sample))

# Merge data
sample_df <-
  info_df %>% 
    dplyr::left_join(run_df, by = c("sample_uuid")) %>% 
    dplyr::left_join(files_df, by = c("npg_run_id")) %>% 
    dplyr::mutate(join_col = dplyr::if_else(project == "6588",sample_donor_id,npg_run_id)) %>% 
    dplyr::left_join(epi_df, by = c("join_col")) %>% 
    dplyr::mutate(Isolate = dplyr::if_else(project == "6310",name,Isolate)) %>% 
    dplyr::select(-join_col) %>% 
    dplyr::filter(!is.na(Country))

```

## Identify species with Kraken and Bracken

Analysis of the short reads used the `qc-short-read` pipeline (revision 103cba2a10; v1.0.2), which depends upon:

-   fastqc (v0.12.1)

-   multiqc (v1.19)

-   kraken2 (v2.1.3)

-   bracken (v2.8)

The results are stored for projects [6310](./pipeline_info/execution_report_2025-09-14_08-15-34.html) and [6588](./pipeline_info/execution_report_2025-09-14_08-18-09.html).

```{bash, eval = FALSE}

module load bsub.py
module load qc-short-read
cd 6310_qc
echo -en "ID\tR1\tR2\n" > 6310.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
bsub.py -q oversubscribed 4 6310_qc_rerun qc-short-read --manifest_of_reads 6310.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false
cd ../6588_qc
echo -en "ID\tR1\tR2\n" > 6588.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub.py -q oversubscribed 4 6588_qc_rerun qc-short-read --manifest_of_reads 6588.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false

```

## Assemble reads with SPAdes and unicycler

Analysis of the short reads used the `assembly-unicycler-short-read` pipeline, developed from the bacass pipeline (v2.5.0; <https://zenodo.org/records/17378847>), which depends upon:

-   QUAST (v5.0.2)

-   Unicycler (v0.5.1)

-   SPAdes (v4.0.0)

The results are stored for projects [6310](./pipeline_info/execution_report_2025-09-11_17-26-05.html) and [6588](./pipeline_info/execution_report_2025-09-11_20-15-21.html).

```{bash, eval = FALSE}

git clone --recurse-submodules git@gitlab.internal.sanger.ac.uk:sanger-pathogens/pipelines/assembly_unicycler_short_read.git
cd assembly_unicycler_short_read
echo -en "ID,R1,R2\n" > 6310.manifest; for f in ../6310_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
echo -en "ID,R1,R2\n" > 6588.manifest; for f in ../6588_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub -o 6310_output.o -e 6310_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6310.manifest --outdir 6310_assemblies
bsub -o 6588_output.o -e 6588_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6588.manifest --outdir 6588_assemblies

```

## Annotate assemblies with Bakta

The assemblies were annotated using Bakta (version 1.11.0).

```{bash, eval = FALSE}

module load bakta/1.11.0
declare -i N; N=1; for f in *.fa; do echo "bakta --db /data/pam/software/bakta/v6.0 --prefix ${f%.fa} --output ${f%.fa}_annotation --gram - --threads 4 ${f}" > bacta_job.${N}; chmod +x ./bacta_job.${N}; N=($N+1); done
bsub -n 4 -M 4000 -R 'select[mem>4000] rusage[mem=4000] span[hosts=1]' -o bakta.%I.o -e bakta.%I.e -J bakta_run[1-4473] ./bacta_job.\${LSB_JOBINDEX}

```

## Evaluate assemblies with Quast

Assembly statistics were calculated with Quast (version 5.0.2).

```{r Parse assembly statistics data}

# Parse assembly statistics
assembly_stats_df <-
  read.table(file = "./sample_data/NP_quast_report.tsv",
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::rename(npg_run_id = Assembly)

# Join to sample information
assembly_stats_df %<>%
  dplyr::left_join(sample_df,
                   by = "npg_run_id")

# Plot and filter


```

### Merge data of two files of sample_df and assembly_stats_df

We use these two files to merge getting chronological data.

```{r, merge data, echo=TRUE}

merge_df <- merge(assembly_stats_df, sample_df, by.x = "filename", by.y = "assembly_file")

View(merge_df)

right_df <- right_join(as_tibble(assembly_stats_df), as_tibble(sample_df), by = c("filename" = "assembly_file"))

View(right_df)

```

### Plot assembly data of baterial genome assemblies between total length and N50 to get quality of genome

We used total_length and N50 of each genomes to find quality of genome with baseline threshold of N50 (y-axis) is 100000bp and total length (x-axis) which is from 1.75Mbp to 2.25Mbp.

```{r, plot-genome assembly, include=TRUE}
ggplot(data = right_df) + geom_point(aes(x = total_length, y = N50, colour = Species)) + geom_hline(yintercept = 100000) + geom_vline(xintercept = c(1.75e+06, 2.25e+06))
```

### Quality controle of merging data

We need to eliminate a few results from sequencing to get the better quality control of sequencing genome

```{r, filter-data, include=TRUE}
summaried_df <- merge_df %>%
  filter(N50 > 100000 & (total_length > 1.75e+06 & total_length < 2.25e+06))
ggplot(data = summaried_df) + geom_point(aes(x = total_length, y = N50, colour = Species)) + geom_hline(yintercept = 100000) + geom_vline(xintercept = c(1.75e+06, 2.25e+06))
```

### Plot graph of identifing bacterial burden in each three countries

```{r, plot-graph, include=TRUE}
ggplot(data = merge_df) + geom_bar(aes(x = Year, fill = Country, position = "dodge")) + facet_wrap(~ Species) + xlab("Year of investigation") + ylab("Number of enrollment\nIn local hospital") + labs(fill = "Study Enrollment\nIn Each Countries", shape = "Species", caption = "Antimicrobial Resistance Study") + scale_shape_manual(values = c(17,19), labels = c("Haemophilus influenzae", "Moraxella catarrhalis")) + coord_cartesian(xlim = c(2004, 2021)) + theme_minimal()
```

## Identify species in datasets

```{r Parse bracken outputs}

parse_bracken <- function(fn) {
  
  read.table(file = gzfile(fn),
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::filter(grepl("s__",Classification)
  
}

species_abundance_df <-
  parse_bracken("./sample_data/6310_bracken_summary_report.tsv.gz") %>%  
  dplyr::full_join(
    parse_bracken("./sample_data/6588_bracken_summary_report.tsv.gz"),
    by = "Classification"
  ) %>%
  dplyr::mutate(Classification = gsub("(.)*\\|s__","",Classification)) %>%
  tidyr::pivot_longer(cols = -Classification, names_to = "npg_run_id", values_to = "matches") %>% 
  dplyr::rename(Species = Classification) %>% 
  dplyr::mutate(matches = dplyr::if_else(is.na(matches),0,matches)) %>% 
  dplyr::group_by(Species) %>% 
  dplyr::mutate(Total = sum(matches)) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(Total > species_match_threshold)

# Join with sample information
species_sample_df <-
  species_abundance_df %>% 
    dplyr::full_join(
      sample_df,
      by = "npg_run_id"
    )
  

# Plot species matches
ggplot(species_abundance_df,
       aes(x = Species,
           y = npg_run_id,
           fill = matches+0.5)) +
  geom_tile() +
  scale_fill_continuous(trans = "log10") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

## Use AMR finder to identify resistance genes

## Use capsule typer to identify serotypes

## Use PopPUNK to categorise isolates

## Use Panaroo to identify genes

## Compare strain frequncies between samples

## Compare gene frequencies between samples
