---
title: "NP AMR analyses"
author: "Nicholas Croucher"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(magrittr)
require(tidyverse)

```

``` {bash, eval=FALSE}

module load PaM/environment
module load shelf
module load bsub.py
module load irods_extractor
module load qc-short-read
iinit
export LSB_DEFAULT_USERGROUP=team284
cd /data/pam/team284/nc3/scratch/NP_AMR/data/
# QC project 6310
cd /data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc
bsub.py --queue oversubscribed --threads 1 4 6310_QC qc-short-read --studyid 6310 --kraken2_db /data/pam/software/kraken2/standard/ --read_len 151 --classification_level S -ansi-log false --kraken2_threads 16 --bracken_threads 16

rm -rf results work *_QC.[oe]; bsub.py --queue oversubscribed --threads 1 4 6310_QC qc-short-read --studyid 6310 --kraken2_db /data/pam/software/kraken2/standard/ --read_len 151 --classification_level S -ansi-log false --kraken2_threads 16 --bracken_threads 16 --type cram; watch bjobs



bsub -o output.o -e error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --input ./test_data/inputs/test_manifest.csv --read_len 150 --threshold 1 --classification_level 'G' --kraken2_threads 10 --outdir my_output

bsub.py 16 6588_QC qc-short-read --studyid 6588 --kraken2_db /data/pam/software/kraken2/standard/ --read_len 150 --classification_level S -ansi-log false

shelf get file -q import.sequencescape_study_id=6310 -q file.file_type=assembly

module load assembly-unicycler-short-read



```

Run assemblies:

``` {bash, eval=FALSE}

module load PaM/environment
module load shelf
export LSB_DEFAULT_USERGROUP=team284
cd /data/pam/team284/nc3/scratch/NP_AMR/data/

```

## Load data from project

Import sample data from project

```{r Process sample data}

parse_sample_file <- function(fn) {
  
  read.table(file = fn,
             sep = "\t",
             comment.char = "",
             header = TRUE)
  
}

process_annotation_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "annotation") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",path)) %>% 
    dplyr::mutate(annotation_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_',npg_run_id)) %>% 
    dplyr::mutate(npg_run_id = gsub(".gff","",npg_run_id)) %>% 
    dplyr::select(-path)
  
}

process_assembly_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "assembly") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/velvet_assembly/contigs.fa","",gsub("/spades_assembly/contigs.fa","",path)))) %>% 
    dplyr::mutate(assembly_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_contigs.fa')) %>% 
    dplyr::select(npg_run_id,assembly_file)
  
}

process_qc_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "qc") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/kraken.report","",path))) %>% 
    dplyr::mutate(qc_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_kraken.report')) %>% 
    dplyr::select(npg_run_id,qc_file)
  
}

# Parse sample files
info_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_info.tsv") %>% 
      dplyr::filter(!grepl("HiC_RMV",sample_supplier_name)) %>%
      dplyr::mutate(project = "6588"),
    parse_sample_file("./sample_data/6310_info.tsv") %>%
      dplyr::mutate(project = "6310")
  )

run_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_runs.tsv") %>% 
      dplyr::select(-name),
    parse_sample_file("./sample_data/6310_runs.tsv") %>% 
      dplyr::select(-name)
  )

fn_6588 <- "./sample_data/6588_files.tsv"
fn_6310 <- "./sample_data/6310_files.tsv"
files_df <-
  dplyr::bind_rows(
    parse_sample_file(fn_6588) %>% process_annotation_file_names(),
    parse_sample_file(fn_6310) %>% process_annotation_file_names()
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_assembly_file_names(),
      parse_sample_file(fn_6310) %>% process_assembly_file_names()
    ),
    by = c("npg_run_id")
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_qc_file_names(),
      parse_sample_file(fn_6310) %>% process_qc_file_names()
    ),
    by = c("npg_run_id")
  )

epi_df <-
  dplyr::bind_rows(
    read.csv("./sample_data/Asian_isolate_data.csv") %>% 
      dplyr::mutate(Description = gsub(" from nasopharyngeal swab","",Description)) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Species = Description) %>% 
      dplyr::mutate(Description = dplyr::if_else(Species == "Mixed bacteria","Mixed bacteria","Single isolate")),
    read.csv("./sample_data/Hampshire_isolate_data.csv") %>% 
      dplyr::mutate(Cohort = Country) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Description = "Single isolate"),
    read.csv("./sample_data/Global-Hinf-metadata-microreact.csv") %>% 
      dplyr::filter(disease == "") %>% 
      dplyr::select(id,country,Year) %>% 
      dplyr::filter(country == "Maela") %>% 
      dplyr::rename(Isolate = id,
                    Region = country) %>% 
      dplyr::mutate(Sample = Isolate,
                    Region = "Maela",
                    Cohort = "Maela",
                    Country = "Thailand",
                    Species = "Haemophilus influenzae",
                    project = "6310",
                    Year = as.numeric(Year)) %>% 
      dplyr::mutate(npg_run_id = gsub("(_.*?)_", "\\1#", Sample)) %>% 
      dplyr::mutate(Description = "Single isolate")
  ) %>% 
  dplyr::mutate(join_col = dplyr::if_else(project == "6588",Sample,npg_run_id)) %>% 
  dplyr::select(-c(project,npg_run_id,Sample))

# Merge data
sample_df <-
  info_df %>% 
    dplyr::left_join(run_df, by = c("sample_uuid")) %>% 
    dplyr::left_join(files_df, by = c("npg_run_id")) %>% 
    dplyr::mutate(join_col = dplyr::if_else(project == "6588",sample_donor_id,npg_run_id)) %>% 
    dplyr::left_join(epi_df, by = c("join_col")) %>% 
    dplyr::mutate(Isolate = dplyr::if_else(project == "6310",name,Isolate)) %>% 
    dplyr::select(-join_col) %>% 
    dplyr::filter(!is.na(Country))

```

## Identify species with Kraken and Bracken

Identification of species from short read data used the qc-short-read pipeline (revision 103cba2a10), which used Kraken2 (version 2.1.3) and Bracken (version 2.8). Pipeline report is [execution_report_2025-09-14_08-15-34.html](./pipeline_info/execution_report_2025-09-14_08-15-34.html).

``` {bash, eval = FALSE}

module load bsub.py
module load qc-short-read
cd 6310_qc
echo -en "ID\tR1\tR2\n" > 6310.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
bsub.py -q oversubscribed 4 6310_qc_rerun qc-short-read --manifest_of_reads 6310.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false
cd ../6588_qc
echo -en "ID\tR1\tR2\n" > 6588.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub.py -q oversubscribed 4 6588_qc_rerun qc-short-read --manifest_of_reads 6588.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false

```

## Assemble reads with SPAdes and unicycler

Assembly used the assembly_unicycler_short_read pipeline (version 25.04.0), which used SPAdes (version 4.0.0) and unicycler (version 0.5.1). Pipeline reports are [execution_report_2025-09-11_20-15-21.html](./pipeline_info/execution_report_2025-09-11_20-15-21.html) and [execution_report_2025-09-11_17-22-43.html](./pipeline_info/execution_report_2025-09-11_17-22-43.html).

``` {bash, eval = FALSE}

git clone --recurse-submodules git@gitlab.internal.sanger.ac.uk:sanger-pathogens/pipelines/assembly_unicycler_short_read.git
cd assembly_unicycler_short_read
echo -en "ID,R1,R2\n" > 6310.manifest; for f in ../6310_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
echo -en "ID,R1,R2\n" > 6588.manifest; for f in ../6588_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub -o 6310_output.o -e 6310_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6310.manifest --outdir 6310_assemblies
bsub -o 6588_output.o -e 6588_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6588.manifest --outdir 6588_assemblies

```

## Annotate assemblies with Bakta

The assemblies were annotated using Bakta (version 1.11).

``` {bash, eval = FALSE}

module load bakta/1.11.0
declare -i N; N=1; for f in *.fa; do echo "bakta --db /data/pam/software/bakta/v6.0 --prefix ${f%.fa} --output ${f%.fa}_annotation --gram - --threads 4 ${f}" > bacta_job.${N}; chmod +x ./bacta_job.${N}; N=($N+1); done
bsub -n 4 -M 4000 -R 'select[mem>4000] rusage[mem=4000] span[hosts=1]' -o bakta.%I.o -e bakta.%I.e -J bakta_run[1-4473] ./bacta_job.\${LSB_JOBINDEX}

```

## Evaluate assemblies with Quast

Assembly statistics were calculated with Quast (version 5.0.2).

``` {r Parse assembly statistics data}

# Parse assembly statistics
assembly_stats_df <-
  read.table(file = "./sample_data/NP_quast_report.tsv",
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::rename(npg_run_id = Assembly)

# Join to sample information
assembly_stats_df %<>%
  dplyr::left_join(sample_df,
                   by = "npg_run_id")

# Plot and filter


```
### Merge data of two files of sample_df and assembly_stats_df
We use these two files to merge getting chronological data.

```{r, merge data, echo=TRUE}

View(assembly_stats_df)
right_df <- right_join(as_tibble(assembly_stats_df), as_tibble(sample_df), by = "assembly_file")

```

### Plot assembly data of baterial genome assemblies between total length and N50 to get quality of genome
We used total_length and N50 of each genomes to find quality of genome with baseline threshold of N50 (y-axis) is 100000bp and total length (x-axis) which is from 1.75Mbp to 2.25Mbp.

``` {r, plot-genome assembly, include=TRUE}

ggplot(data = assembly_stats_df) + geom_point(aes(x = `Total length`, y = N50, colour = Species)) + geom_hline(yintercept = 100000) + scale_x_log10() + geom_vline(xintercept = c(1.75e+06, 2.25e+06))
```
### Quality controle of merging data
We need to eliminate a few results from sequencing to get the better quality control of sequencing genome

``` {r, filter-data, include=TRUE}
summaried_df <- assembly_stats_df %>%
  filter(N50 > 10000 & (`Total length` > 1.5e+06 & `Total length` < 2.5e+06))
ggplot(data = summaried_df) + geom_point(aes(x = `Total length`, y = N50, colour = Species)) + geom_hline(yintercept = 80000) + geom_vline(xintercept = c(1.725e+06, 2.15e+06))

```
### Plot graph of identifing bacterial burden in each three countries

``` {r, plot-graph, include=TRUE}
ggplot(data = assembly_stats_df) + geom_bar(aes(x = Year, fill = Country, position = "dodge")) + facet_wrap(~ Species) + xlab("Year of investigation") + ylab("Number of enrollment\nIn local hospital") + labs(fill = "Study Enrollment\nIn Each Countries", shape = "Species", caption = "Antimicrobial Resistance Study") + scale_shape_manual(values = c(17,19), labels = c("Haemophilus influenzae", "Moraxella catarrhalis")) + coord_cartesian(xlim = c(2004, 2021)) + theme_minimal()
```

## Identify species in datasets

``` {r Parse bracken outputs}

parse_bracken <- function(fn) {
  
  read.table(file = gzfile(fn),
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::filter(grepl("s__",Classification))
  
}
species_match_threshold = 10000000
species_abundance_df <-
  parse_bracken("./sample_data/6310_bracken_summary_report.tsv.gz") %>%  
  dplyr::full_join(
    parse_bracken("./sample_data/6588_bracken_summary_report.tsv.gz"),
    by = "Classification"
  ) %>%
  dplyr::mutate(Classification = gsub("(.)*\\|s__","",Classification)) %>%
  tidyr::pivot_longer(cols = -Classification, names_to = "npg_run_id", values_to = "matches") %>% 
  dplyr::rename(Species = Classification) %>% 
  dplyr::mutate(matches = dplyr::if_else(is.na(matches),0,matches)) %>% 
  dplyr::group_by(Species) %>% 
  dplyr::mutate(Total = sum(matches)) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(Total > species_match_threshold)

# Join with sample information
species_sample_df <-
  species_abundance_df %>% 
    dplyr::full_join(
      sample_df,
      by = "npg_run_id"
    )
  

# Plot species matches
ggplot(species_sample_df,
       aes(x = Species.x,
           y = npg_run_id,
           fill = matches+0.5)) +
  geom_tile() + facet_grid(Species.y ~., scales = "free_y") + scale_fill_continuous(trans = "log10") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
## We had filted dataset to calculate proportion and matching proportion for seeing how proportion in dataset is the right matching species

```{r create a proportion}

df <- species_sample_df %>%
  dplyr::mutate(Species.x = gsub("_"," ", Species.x)) %>%
  group_by(npg_run_id) %>%
  mutate(proportion = matches / sum(matches)) %>%
dplyr::mutate(Matching_proportion = dplyr::if_else (Species.x == Species.y, proportion, NA_real_)) %>%
tidyr::fill(Matching_proportion, .direction = "updown")
df
```
## We plot matching proportion for each types of species (Moraxella catarrhalis, Haemophilus influenzae and Mixed bacteria) to get the high quality species 

```{r plot proportion}

filtered_df <- df %>%
  filter(Species.y %in% c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") & Species.y == Species.x)
summary_df <- filtered_df %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))
ggplot(filtered_df, aes(x = Species.y, y = proportion)) + geom_violin()+ geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") + labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") + theme_minimal()

```
## We try to remove the species with lower proportion of 0.95

```{r filter the proportion }

filtered_df <- df %>%
  
  filter(Species.y %in% c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") & Species.y == Species.x & proportion > 0.95)
summary_df <- filtered_df %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))
ggplot(filtered_df, aes(x = Species.y, y = proportion)) + geom_violin()+ geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") + labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") + theme_minimal()
```
## plot the characteristic of the filter data (country and years)

```{r plot the filter dataset}

plot_column <- df %>%
  filter(Species.y %in% c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria"),
         Species.y == Species.x,
         proportion > 0.95) %>%
  select(Country, Year, Species.y, proportion, Matching_proportion, Total, npg_run_id, matches) %>%
  distinct()
plot_column

ggplot(data = plot_column) + geom_bar(aes(x = Year, fill = Country)) + facet_wrap(~ Species.y) + labs(title = "Total by Country and Years", x = "Years", y = "Total", fill = "Countries") + theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r rename Global_Hinfl dataset}
selected_Global_data <- Global_Hinf_metadata_microreact[, grep("^(id|Year)$", names(Global_Hinf_metadata_microreact))]
selected_Global_data
rename_Global_data <- selected_Global_data %>%
  select(id, Year) %>%
  rename(Isolate = id, new_Year = Year)
rename_Global_data
```
```{r merge both datasets of Global and epi}

join_epi_Global_data <- left_join(rename_Global_data, epi_df, by = "Isolate") %>%
  mutate(Year = if_else(!is.na(new_Year), new_Year, Year)) %>%
  select(-new_Year)
join_epi_Global_data

```

## We want to keep sampels with threshold:
N50 above 80,000
Total length above 1,725,000 bp
Total length below 2,150,000 bp
and Matching proportion > 0.95 of *Moraxella catarrhalis* and *Haemophilus influenzae* colony

## Tidy up the column name of dataset 
Species.x to orginal species
Species.y to matching species


## Use AMR finder to identify resistance genes

## Use capsule typer to identify serotypes

## Use PopPUNK to categorise isolates

## Use Panaroo to identify genes

## Compare strain frequncies between samples

## Compare gene frequencies between samples

