---
title: "NP AMR analyses"
author: "Nicholas Croucher"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(magrittr)
require(tidyverse)
require(stringr)

```

## Load data from project

Import sample data from project

```{r Process sample data}

parse_sample_file <- function(fn) {
  
  read.table(file = fn,
             sep = "\t",
             comment.char = "",
             header = TRUE)
  
}

process_annotation_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "annotation") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",path)) %>% 
    dplyr::mutate(annotation_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_',npg_run_id)) %>% 
    dplyr::mutate(npg_run_id = gsub(".gff","",npg_run_id)) %>% 
    dplyr::select(-path)
  
}

process_assembly_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "assembly") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/velvet_assembly/contigs.fa","",gsub("/spades_assembly/contigs.fa","",path)))) %>% 
    dplyr::mutate(assembly_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_contigs.fa')) %>% 
    dplyr::select(npg_run_id,assembly_file)
  
}

process_qc_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "qc") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/kraken.report","",path))) %>% 
    dplyr::mutate(qc_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_kraken.report')) %>% 
    dplyr::select(npg_run_id,qc_file)
  
}

# Parse sample files
drak_info_df <-
  read.csv("./sample_data/Drakenstein_epi.csv") %>% 
      dplyr::mutate(Year = as.numeric(Year)) %>% 
      dplyr::left_join(
        read_csv("./sample_data/Drakenstein_accessions.csv"),
        by = c("Isolate"="run_accession")
      )

info_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_info.tsv") %>% 
      dplyr::filter(!grepl("HiC_RMV",sample_supplier_name)) %>%
      dplyr::mutate(project = "6588"),
    parse_sample_file("./sample_data/6310_info.tsv") %>%
      dplyr::mutate(project = "6310"),
    drak_info_df %>% 
      dplyr::mutate(name = Isolate,
                    sample_accession_number = sample_accession,
                    sample_common_name = "Moraxella catarrhalis",
                    project = "Drakenstein",
                    sample_supplier_name = Isolate
      ) %>% 
      dplyr::select(name,sample_accession_number,sample_common_name,project,sample_supplier_name)
  )

run_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_runs.tsv") %>% 
      dplyr::select(-name),
    parse_sample_file("./sample_data/6310_runs.tsv") %>% 
      dplyr::select(-name)
  )

fn_6588 <- "./sample_data/6588_files.tsv"
fn_6310 <- "./sample_data/6310_files.tsv"
files_df <-
  dplyr::bind_rows(
    parse_sample_file(fn_6588) %>% process_annotation_file_names(),
    parse_sample_file(fn_6310) %>% process_annotation_file_names()
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_assembly_file_names(),
      parse_sample_file(fn_6310) %>% process_assembly_file_names()
    ),
    by = c("npg_run_id")
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_qc_file_names(),
      parse_sample_file(fn_6310) %>% process_qc_file_names()
    ),
    by = c("npg_run_id")
  )

epi_df <-
  dplyr::bind_rows(
    read.csv("./sample_data/Asian_isolate_data.csv") %>% 
      dplyr::mutate(Description = gsub(" from nasopharyngeal swab","",Description)) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Species = Description) %>% 
      dplyr::mutate(Description = dplyr::if_else(Species == "Mixed bacteria","Mixed bacteria","Single isolate")),
    read.csv("./sample_data/Hampshire_isolate_data.csv") %>% 
      dplyr::mutate(Cohort = Country) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Description = "Single isolate"),
    read.csv("./sample_data/Global-Hinf-metadata-microreact.csv") %>% 
      dplyr::filter(disease == "") %>% 
      dplyr::select(id,country,Year) %>% 
      dplyr::filter(country == "Maela") %>% 
      dplyr::rename(Isolate = id,
                    Region = country) %>% 
      dplyr::mutate(Sample = Isolate,
                    Region = "Maela",
                    Cohort = "Maela",
                    Country = "Thailand",
                    Species = "Haemophilus influenzae",
                    project = "6310",
                    Year = as.numeric(Year)) %>% 
      dplyr::mutate(npg_run_id = gsub("(_.*?)_", "\\1#", Sample)) %>% 
      dplyr::mutate(Description = "Single isolate"),
    drak_info_df %>% 
      dplyr::mutate(Isolate = sample_accession) %>% 
      dplyr::select(-sample_accession) %>% 
      dplyr::mutate(Description = "Single isolate")
  ) %>% 
  dplyr::mutate(join_col = dplyr::case_when(
    project == "6588" ~ Sample,
    Country == "South Africa" ~ Isolate,
    TRUE ~ npg_run_id)) %>% 
  dplyr::select(-c(project,npg_run_id,Sample))

# Merge data
sample_df <-
  info_df %>% 
    dplyr::left_join(run_df, by = c("sample_uuid")) %>% 
    dplyr::left_join(files_df, by = c("npg_run_id")) %>% 
    dplyr::mutate(npg_run_id = dplyr::if_else(project == "Drakenstein",
                                              sample_accession_number,
                                              npg_run_id)) %>%
    dplyr::mutate(join_col = dplyr::if_else(project == "6588",sample_donor_id,npg_run_id)) %>% 
    dplyr::left_join(epi_df, by = c("join_col")) %>% 
    dplyr::mutate(Isolate = dplyr::if_else(project == "6310",name,Isolate)) %>% 
    dplyr::select(-join_col) %>% 
    dplyr::filter(!is.na(Country))
sample_df

```

## Identify species with Kraken and Bracken

Analysis of the short reads used the `qc-short-read` pipeline (revision 103cba2a10; v1.0.2), which depends upon:

-   fastqc (v0.12.1)

-   multiqc (v1.19)

-   kraken2 (v2.1.3)

-   bracken (v2.8)

The results are stored for projects [6310](./pipeline_info/execution_report_2025-09-14_08-15-34.html) and [6588](./pipeline_info/execution_report_2025-09-14_08-18-09.html).

```{bash, eval = FALSE}

module load bsub.py
module load qc-short-read
cd 6310_qc
echo -en "ID\tR1\tR2\n" > 6310.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
bsub.py -q oversubscribed 4 6310_qc_rerun qc-short-read --manifest_of_reads 6310.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false
cd ../6588_qc
echo -en "ID\tR1\tR2\n" > 6588.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub.py -q oversubscribed 4 6588_qc_rerun qc-short-read --manifest_of_reads 6588.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false

```

Download data from the Drachenstein study; results are available [here](./pipeline_info/execution_report_2025-10-22_06-07-18.html):

``` {bash, eval = FALSE}

cd DRAK_study
module load PaM/environment
module load qc-short-read
module load bsub.py

bsub.py -q oversubscribed 4 DRAK_qc qc-short-read --manifest_ena DRAK.manifest --accession_type run --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false

```

## Assemble reads with SPAdes and unicycler

Analysis of the short reads used the `assembly-unicycler-short-read` pipeline (v1.2.5; commit 92464961e453d624e88e69da06670638fb90ede0), developed from the bacass pipeline (v2.5.0; <https://zenodo.org/records/17378847>), which depends upon:

-   QUAST (v5.0.2)

-   Unicycler (v0.5.1)

-   SPAdes (v4.0.0)

The results are stored for projects [6310](./pipeline_info/execution_report_2025-09-11_17-26-05.html) and [6588](./pipeline_info/execution_report_2025-09-11_20-15-21.html).

```{bash, eval = FALSE}

git clone --recurse-submodules git@gitlab.internal.sanger.ac.uk:sanger-pathogens/pipelines/assembly_unicycler_short_read.git
cd assembly_unicycler_short_read
echo -en "ID,R1,R2\n" > 6310.manifest; for f in ../6310_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
echo -en "ID,R1,R2\n" > 6588.manifest; for f in ../6588_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub -o 6310_output.o -e 6310_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6310.manifest --outdir 6310_assemblies
bsub -o 6588_output.o -e 6588_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6588.manifest --outdir 6588_assemblies

```

Assemble data from the Drachenstein study; results are available [here](./pipeline_info/execution_report_2025-10-22_12-50-46.html):

``` {bash, eval = FALSE}

cd DRAK_study
module load PaM/environment
module load qc-short-read
module load bsub.py

bsub -o drak_output.o -e drak_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest_ena /data/pam/team284/nc3/scratch/NP_AMR/data/DRAK_study/DRAK.manifest --accession_type run --outdir /data/pam/team284/nc3/scratch/NP_AMR/data/DRAK_study/assemblies

```

## Annotate assemblies with Bakta

The assemblies were annotated using Bakta (version 1.11.0).

```{bash, eval = FALSE}

module load bakta/1.11.0
declare -i N; N=1; for f in *.fa; do echo "bakta --db /data/pam/software/bakta/v6.0 --prefix ${f%.fa} --output ${f%.fa}_annotation --gram - --threads 4 ${f}" > bacta_job.${N}; chmod +x ./bacta_job.${N}; N=($N+1); done
bsub -n 4 -M 4000 -R 'select[mem>4000] rusage[mem=4000] span[hosts=1]' -o bakta.%I.o -e bakta.%I.e -J bakta_run[1-4473] ./bacta_job.\${LSB_JOBINDEX}

```

This was also run on the Drachenstein isolates, after 1427/1429 isolates were successfully downloaded and assembled:

```{bash, eval = FALSE}

module load bakta/1.11
declare -i N; N=1; for f in *.fa; do echo "bakta --db /data/pam/software/bakta/v6.0 --prefix ${f%.fa} --output ${f%.fa}_annotation --gram - --threads 4 ${f}" > bacta_job.${N}; chmod +x ./bacta_job.${N}; N=($N+1); done
bsub -n 4 -M 12000 -R 'select[mem>12000] rusage[mem=12000] span[hosts=1]' -o bakta.%I.o -e bakta.%I.e -J bakta_run[1-1427] ./bacta_job.\${LSB_JOBINDEX}

```

## Evaluate assemblies with Quast

Assembly statistics were calculated with Quast (version 5.0.2).

```{r Parse assembly statistics data}

# Parse assembly statistics
assembly_stats_df <-
  dplyr::bind_rows(
    read.table(file = "./sample_data/NP_quast_report.tsv",
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE),
    read.table(file = "./sample_data/Drakenstein_summary_quast_report.tsv",
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
      dplyr::mutate(Assembly = gsub(".assembly","",Assembly))
  ) %>% 
    dplyr::rename(npg_run_id = Assembly)
# Join to sample information
assembly_stats_df %>%
  dplyr::left_join(sample_df,
                   by = "npg_run_id")
assembly_stats_df
```

### Merge two datasets of sample_df and assembly_stats_df

We will use these two files to merge getting chronological data set.

```{r, merge data, echo=TRUE}
assembly_join_sample <- 
  right_join(as_tibble(assembly_stats_df),
             as_tibble(sample_df),
             by = "npg_run_id")
assembly_join_sample

minimum_length <- 1750000
maximum_length <- 2250000
n50_threshold <- 100000

```

### Plot assembly data of baterial genomic assemblies between total length and N50 to get quality of genomesz
We used total length and N50 of each genomes to find quality of genomes with baseline threshold of N50 (y-axis) is `r n50_threshold`` bp and total length (x-axis) which is from `r minimum_length`` Mbp to `r maximum_length` Mbp.

```{r, plot-genome assembly, include=TRUE}

ggplot(data = assembly_join_sample) + 
  geom_point(aes
             (x = `Total length`, 
               y = N50, 
               colour = Species)) +
  geom_hline(yintercept = n50_threshold) + 
  scale_x_log10() + 
  geom_vline(xintercept = c(minimum_length, 
                            maximum_length))
                                            
```

### We need to plot of merging dataset to be known quality control of each species and knowing limitted threshold to be accepted in each species.
We need to eliminate a few results from sequencing to get the better quality control of each species.

```{r, filter-data, include=TRUE}

summary_df <- 
  assembly_join_sample %>%
  filter(N50 > n50_threshold & 
           (`Total length` > minimum_length & 
              `Total length` < maximum_length))
summary_df

ggplot(data = summary_df) + 
  geom_point(aes
             (x = `Total length`, 
               y = N50, colour = Species)) + 
  geom_hline(yintercept = n50_threshold) + 
  geom_vline(xintercept = c(minimum_length, maximum_length))

```

### Plot graph of identifing bacterial burden in each three countries including South Africa

```{r, plot-graph, include=TRUE}

ggplot(data = assembly_join_sample) + 
  geom_bar(aes
           (x = Year, 
             fill = Country), 
           position = "dodge", 
           stat = "count") + 
  facet_wrap(~ Species) + 
  xlab("Year of investigation") + 
  ylab("Number of enrollment\nIn local hospital") + 
  labs(fill = "Study Enrollment\nIn Each Countries", 
       shape = "Species", 
       caption = "Antimicrobial Resistance Study") + 
  scale_shape_manual(values = c(17,19, 20), 
                     labels = c("Haemophilus influenzae", "Moraxella catarrhalis", "Mixed bacteria")) + coord_cartesian(xlim = c(2005, 2021)) + 
  theme_minimal()
```

```{r filter npg_run_id list of Moraxella cattarrhalis}
filter_Moraxella <- summary_df %>%
  filter(Species == "Moraxella catarrhalis") %>%
select(npg_run_id) %>%
distinct()
filter_Moraxella
```

```{r generate file of npg_run_id Moraxella cattarrhalis}
write.table(filter_Moraxella,
            file = "./sample_data/npg_run_id_Morax2_SA.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r filter npg_run_id of Haemophilus influenzae}
filter_Haemophilus <- summary_df %>%
  filter(Species == "Haemophilus influenzae") %>%
select(npg_run_id) %>%
  distinct()
filter_Haemophilus

```

```{r generate npg_run_id file of Haemophilus influenzae}
write.table(filter_Haemophilus,
            file = "./sample_data/npg_run_id_Haemo2_SA.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

## Identify species in datasets

```{r Parse bracken outputs}

parse_bracken <- function(fn) {
  
  read.table(file = gzfile(fn),
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::filter(grepl("s__",Classification))
}

species_match_threshold = 10000000

species_abundance_df <-
  parse_bracken("./sample_data/6310_bracken_summary_report.tsv.gz") %>%  
  dplyr::full_join(
    parse_bracken("./sample_data/6588_bracken_summary_report.tsv.gz"),
    by = "Classification"
  ) %>%
  dplyr::full_join(parse_bracken("./sample_data/Drakenstein_bracken_summary_report.tsv.gz"),
    by = "Classification"
  ) %>%
  dplyr::mutate(Classification = gsub("(.)*\\|s__","",Classification)) %>%
  tidyr::pivot_longer(cols = -Classification, names_to = "npg_run_id", values_to = "matches") %>% 
  dplyr::rename(Species = Classification) %>% 
  dplyr::mutate(matches = dplyr::if_else(is.na(matches),0,matches)) %>% 
  dplyr::group_by(Species) %>% 
  dplyr::mutate(Total = sum(matches)) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(Total > species_match_threshold)

# Join with sample information
species_sample_df <-
  species_abundance_df %>% 
    dplyr::full_join(
      sample_df,
      by = "npg_run_id"
    )

# Plot species matches
ggplot(species_sample_df,
       aes(x = Species.x,
           y = npg_run_id,
           fill = matches+0.5)) +
  geom_tile() +
  facet_grid(Species.y ~., scales = "free_y") +
  scale_fill_continuous(trans = "log10") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```
```{r merge two datasets between assembly_stats_df and species_sample_df}
assembly_species_merge <- left_join(assembly_stats_df, species_sample_df, by = "npg_run_id")
assembly_species_merge

```

```{r filtering assembly_species_merge to get quality genomes using N50 and Total length}
clean_assembly_species <- assembly_species_merge %>%
  filter(!is.na(Species.y),
         !is.na(`Total length`),
         !is.na(N50),
         !is.na(Species.x))
clean_assembly_species

ggplot(data = clean_assembly_species) + 
  geom_point(aes
             (x = `Total length`, 
               y = N50, 
               colour = Species.y)) +
  geom_hline(yintercept = 80000) + 
  scale_x_log10() + 
  geom_vline(xintercept = c(minimum_length, 
                            maximum_length))

```

```{r filtering genomes for getting good genomic quality by N50 and Total length}
assembly_species_filtering <- 
  assembly_species_merge %>%
  filter(N50 > n50_threshold & 
           (`Total length` > minimum_length & 
              `Total length` < maximum_length))
assembly_species_filtering

ggplot(data = assembly_species_filtering) + 
  geom_point(aes
             (x = `Total length`, 
               y = N50, colour = Species.y)) + 
  geom_hline(yintercept = n50_threshold) + 
  geom_vline(xintercept = c(minimum_length, maximum_length))

```

### Re-filter dataset to get specific matching genomes of Moraxella cattarrhalis and Haemophilus influenzae
```{r filter two specific genomics of Mcat and Hinflu}
specific_filter <- assembly_species_filtering %>%
  mutate(Species.x = recode(Species.x, "Moraxella_catarrhalis" = "Moraxella catarrhalis", "Haemophilus_influenzae" = "Haemophilus influenzae")) %>%
  filter(!is.na(Species.x),
         !is.na(Species.y),
         Species.x %in% c("Moraxella catarrhalis", "Haemophilus influenzae"), 
         Species.x == Species.y) %>%
  select(npg_run_id, `Total length`, `GC (%)`, N50, Species.x, matches, Total, name, sample_common_name, total_reads, Isolate, Cohort, Country, Region, Year, Description, Species.y) %>%
  group_by(npg_run_id) %>%
  mutate(proportion = matches / sum(matches)) %>%
  mutate(Matching_proportion = dplyr::if_else (Species.x == Species.y, proportion, NA_real_)) %>%
  fill(Matching_proportion, .direction = "updown")
specific_filter

```


```{r filter Moraxella cattarrhalis specie by npg_run_id}

Morax_filter <- specific_filter %>%
  filter(Species.y == "Moraxella catarrhalis") %>%
select(npg_run_id) %>%
distinct()
Morax_filter
```

```{r generate list of npg_run_id Moraxella cattarrhalis}
write.table(Morax_filter,
            file = "./sample_data/npg_run_id_2ndfinal_Mcattarrhalis.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r filter npg_run_id of Haemophilus influenzae genomes}

Hinflu_filter <- specific_filter %>% 
  filter(Species.y == "Haemophilus influenzae") %>% 
  select(npg_run_id) %>%
  distinct()
Hinflu_filter

```

```{r generate npg_run_id list of Haemophilus influenzae}

write.table(Hinflu_filter,
            file = "./sample_data/npg_run_id_2ndfinal_Hinfluenzae.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)

```




## We had filted dataset to calculate proportion and matching proportion for seeing how proportion in dataset is the right matching species

```{r create a proportion}

proportion_df <- species_sample_df %>%
  filter(!is.na(Species.x), 
         !is.na(Species.y), 
         !is.na(matches)) %>%
  dplyr::mutate(Species.x = gsub("_"," ", Species.x)) %>%
  group_by(npg_run_id) %>%
  mutate(proportion = matches / sum(matches)) %>%
dplyr::mutate(Matching_proportion = dplyr::if_else (Species.x == Species.y, proportion, NA_real_)) %>%
tidyr::fill(Matching_proportion, .direction = "updown") %>%
  ungroup()
proportion_df
proportion_threshold <- 0.95
```

## We plot matching proportion for each types of species (Moraxella catarrhalis, Haemophilus influenzae and Mixed bacteria) to get the high quality species

We will filter at a species matching proportion of `r proportion_threshold`.

```{r plot proportion}

filter_proportion_df <- proportion_df %>%
  filter(!is.na(Species.y), 
         Species.y %in%
           c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") &
           Species.y == Species.x) %>%
  group_by(Species.y, proportion) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))
ggplot(filter_proportion_df, aes(x = Species.y, y = proportion)) +
  geom_violin()+
  geom_hline(yintercept = proportion_threshold, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") +
  theme_minimal()

```

## We use the genus level to analyse Moraxella

```{r Moraxella at genus level}

genus_level_df <-
  proportion_df %>% 
    dplyr::mutate(Observed_genus = stringr::str_extract(Species.x, '[A-Za-z]+'),
                  Expected_genus = stringr::str_extract(Species.y, '[A-Za-z]+'))

filtered_genus_df <-
  genus_level_df %>%
    dplyr::filter(!is.na(Species.y), 
                  Species.y %in% c("Moraxella catarrhalis", "Haemophilus influenzae")) %>% 
    dplyr::group_by(npg_run_id) %>%
    dplyr::mutate(Total = sum(matches, na.rm = TRUE)) %>% 
    dplyr::ungroup() %>% 
    dplyr::group_by(npg_run_id, Observed_genus) %>%
    dplyr::mutate(proportion = sum(matches, na.rm = TRUE)/Total) %>% 
    dplyr::ungroup() %>%
    dplyr::filter(Expected_genus == Observed_genus)

ggplot(filtered_genus_df, aes(x = Expected_genus, y = proportion)) +
  geom_violin() +
  geom_hline(yintercept = proportion_threshold, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Genus", x = "Species", y = "Mean Proportion") +
  theme_minimal()

```

## We try to remove the species with lower proportion of 0.95
We want to keep samples with threshold:
N50 above 80,000, Total length above 1,725,000 bp and Total length below 2,150,000 bp, matching proportion > 0.95 of *Moraxella catarrhalis* and *Haemophilus influenzae* colony.

```{r filter the proportion }

filtered_proportion <- proportion_df %>%
  filter(!is.na(Species.y), 
         Species.y %in%
           c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") &
           Species.y == Species.x & proportion > proportion_threshold)

concluded_proportion <- filtered_proportion %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))

ggplot(filtered_proportion,
       aes(x = Species.y, y = proportion)) +
  geom_violin()+
  geom_hline(yintercept = proportion_threshold, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") +
  theme_minimal()
```

## plot the characteristic of the filter data (country and years)

```{r plot the filter dataset}

plot_genus_df <-
  filtered_genus_df %>%
  dplyr::filter(!is.na(Species.y), 
                Species.y %in%
                  c("Moraxella catarrhalis", "Haemophilus influenzae") &
                Species.y == Species.x &
                proportion > proportion_threshold) %>%
  dplyr::select(
    npg_run_id,
    Country,
    Year,
    Species.y,
    proportion,
    Matching_proportion,
    Total,
    matches) %>%
  dplyr::distinct()
plot_genus_df

ggplot(data = plot_genus_df) +
  geom_bar(aes(x = Year, fill = Country)) +
  facet_wrap(~ Species.y) +
  labs(
    title = "Total by Countries and Years",
    x = "Years",
    y = "Total",
    fill = "Country"
    ) +
  theme_minimal() +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1)
    )
```

```{r filter npg_run_id of Moraxella }

filter_Moraxella <- plot_genus_df %>%
  filter(Species.y == "Moraxella catarrhalis") %>%
select(npg_run_id) %>%
distinct()
filter_Moraxella
```

```{r generate npg_run_id list of Moraxella}
write.table(filter_Moraxella,
            file = "./sample_data/npg_run_id_Moraxella_cat.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r filter npg_run_id of Haemophilus}
filter_Haemophilus <- plot_genus_df %>%
  filter(Species.y == "Haemophilus influenzae") %>%
select(npg_run_id) %>%
  distinct()
filter_Haemophilus
```

```{r generate npg_run_id list of Haemophilus}
write.table(filter_Haemophilus,
            file = "./sample_data/npg_run_id_Haemo_influ.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

## Calculate genomes of each species and each countries in each years.

```{r calculate genomes of each species}

table(plot_genus_df$Species.y)
table(plot_genus_df$Country)
table(plot_genus_df$Year)
```

```{r Moraxella in each country and each year}

moraxella_Country_Year <-
  plot_genus_df %>%
  filter(Species.y == "Moraxella catarrhalis")

table(moraxella_Country_Year$Country,
      moraxella_Country_Year$Year)

plot_moraxella <-
  moraxella_Country_Year %>%
  count(Country, Year)

ggplot(plot_moraxella,
       aes(x = Year, y = n, fill = Country)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3, size = 3
    ) +
  ggtitle("Distribution of Moraxella catarrhalis\nin each coutries and years")

```

```{r Haemophilus influenzae in each country and each year}
Haemo_Country_Year <-
  plot_genus_df %>%
  filter(Species.y == "Haemophilus influenzae")

table(Haemo_Country_Year$Country, 
      Haemo_Country_Year$Year)
plot_Haemo <- 
  Haemo_Country_Year %>%
  count(Country, Year)
ggplot(
  plot_Haemo, aes(x = Year, y = n, fill = Country)) +
  geom_bar(stat = "identity") +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3, size = 3
    ) +
  ggtitle("Distribution of Haemophilus influenzae\nin each countries and years")

```

```{r get genomes to process}

mcat_list <-
  filtered_genus_df %>% 
    dplyr::filter(Species.x == "Moraxella catarrhalis") %>% 
    dplyr::select(npg_run_id)

write.table(mcat_list,
            file = "./sample_data/Mcat_single_isolates.txt",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r import Global Hinf metadata microreact}
Global_Hinf_metadata_microreact <- read.csv(file = "sample_data/Global-Hinf-metadata-microreact.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

```{r rename Global_Hinf_metadata_microreact dataset}
selected_Global_data <-
  Global_Hinf_metadata_microreact[, grep("^(id|Year)$", 
                                         names(Global_Hinf_metadata_microreact))]
selected_Global_data
rename_Global_data <-
  selected_Global_data %>%
  select(id, Year) %>%
  rename(
    Isolate = id, new_Year = Year
    )
rename_Global_data
```

```{r merge both datasets of Global and epi}

join_epi_Global_data <-
  left_join(
    rename_Global_data, epi_df,
            by = "Isolate"
    ) %>%
  mutate(
    Year = if_else(!is.na(new_Year),
                   new_Year, Year)
    ) %>%
  select(-new_Year)
join_epi_Global_data

```


## Use AMR finder to identify resistance genes

## Use capsule typer to identify serotypes

## Use PopPUNK to categorise isolates


## Use Panaroo to identify genes
After we got the gene presence absence from Moraxella catarrhalis Panaroo, then we extracted that csv file for doing tidy column, calculate the gene frequencies and number of genomes then final plotting scatter graph and plot histograms between countries.

```{r tidy up the Moraxella cattarrhalis gene presence absence columns to longer format columns}

Moraxella_gene_presence <- read.csv(
  "./panaroo_moraxella_file/Moraxella_gene_presence_absence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Moraxella_gene_presence
```

```{r left_join Moraxella gene presence dataset with sample_df}

Moraxella_gene_sample <- 
  left_join(Moraxella_gene_presence, 
            sample_df, 
            by = "npg_run_id")
Moraxella_gene_sample
```

```{r classifying Moraxella gene sample dataset in early and late year of UK country}

UK_early_late_df <- Moraxella_gene_sample %>%
  filter(Country == "United Kingdom",
         Year %in% c(2007:2011, 2019:2020)
         ) %>%
  mutate(Period = 
           case_when(Year >= 2007
                     & Year <= 2011 ~ "Early",
                     Year >= 2019 & Year <= 2020 ~ "Late")
         )
```

```{r calculate the number of Moraxella genome in each dataset using npg_run_id}
Genome_Count <- Moraxella_gene_sample %>%
  filter(Country %in% 
           c("United Kingdom", "Cambodia", "Thailand"),
    Species == "Moraxella catarrhalis", 
    !is.na(npg_run_id)
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id), 
            .groups = "drop")
Genome_Count
```

```{r calculate the number genome count and gene frequencies of Moraxella in dataset using npg_run_id}
Genome_Count <- Moraxella_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Genome_Count

Count_CDS <- Moraxella_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Count_CDS
gene_freq <- Count_CDS %>%
  left_join(Genome_Count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
gene_freq
```

```{r reform gene_freq to wide format}
gene_freq_wide <- gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
  pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
gene_freq_wide
```

```{r plot scatter graph of frequency genes in between United Kingdom and Cambodia Country}

ggplot(gene_freq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "red")+ 
  labs(title = "Moraxella catarrhalis\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph of Moraxella catarrhalis frequency genes in between United Kingdom and Thailand Country}
ggplot(gene_freq_wide, 
       aes(x = Thailand, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "red") + 
  labs(title = "Moraxella catarrhalis Frequency genes by Country", 
       x = "Thailand", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph of Moraxella frequency genes in between Cambodia and Thailand Country}
ggplot(gene_freq_wide, 
       aes(x = Thailand, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "red")+ 
  labs(title = "Moraxella catrrhalis\nFrequency genes by Country", 
       x = "Thailand", 
       y = "Cambodia") + 
  theme_minimal()
```

```{r plot histogram of Moraxella catrrhalis frequency genes comparing with each countries}

ggplot(gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal()

```

### We will calculate genome count and frequency genes of Haemophilus influenzae

```{r import Haemophilus influenzae of gene presece absence csv file for tidy up the column}
Haemophilus_gene_presence <- read.csv("./panaroo_haemophilus_file/Haemo_gene_presence_absence.csv", 
                                      header = TRUE, 
                                      sep = ",", 
                                      stringsAsFactors = FALSE, 
                                      comment.char = "", 
                                      check.names = FALSE) %>%
  tidyr::pivot_longer(cols = 
                        -c(Gene,
                           `Non-unique Gene name`, 
                           Annotation), 
                      names_to = "npg_run_id", 
                      values_to = "CDS")
Haemophilus_gene_presence
```

```{r merge Haemophilus gene presence absence and sample_df with left join function}

Haemo_gene_preab_sample <- left_join(Haemophilus_gene_presence, sample_df, by = "npg_run_id")
Haemo_gene_preab_sample

```

```{r calculate number of genome count and frequency genes of Haemophilus influenzae}
# Count distinct genomes per country
Haemo_gene_count <- Haemo_gene_preab_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Haemo_gene_count
# Count CDS occurrences per gene and country
Haemo_CDS_count <- Haemo_gene_preab_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemo_CDS_count

# Calculate frequency genes
Haemo_gene_freq <- Haemo_CDS_count %>%
  left_join(Haemo_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
View(Haemo_gene_freq)
```

```{r reform Haemo_gene_freq columns to wide format columns}
Haemo_gene_freq_wide <- Haemo_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
```

```{r plot scatter graph of Haemophilus influenzae frequency genes in between United Kingdom and Cambodia}
ggplot(Haemo_gene_freq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "red") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Haemophilus influenzae \nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()

```

```{r plot scatter graph of Haemophilus influenzae frequency genes in between United Kingdom and Thailand Country}
ggplot(Haemo_gene_freq_wide, 
       aes(x = Thailand, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "orange") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Haemophilus influenzae \nFrequency genes by Country", 
       x = "Thailand", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph of Haemophilus influenzae frequency genes in between Cambodia and Thailand}
ggplot(Haemo_gene_freq_wide, 
       aes(x = Thailand, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Haemophilus influenzae \nFrequency genes by Country", 
       x = "Thailand", 
       y = "Cambodia") + 
  theme_minimal()
```

```{r plot histogram of Haemophilus influenzae with frequency genes in each countries}
ggplot(Haemo_gene_freq, 
       aes(x = Frequency)) + 
  geom_histogram() + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Haemophilus influenzae\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```

## We will extract the Moraxella catarrhalis and Haemophilus influenzae files from panaroo modifying scripts are named gene presence and absence csv to calculate the number of genomes and frequnecy genes then plot the scatter graph and histogram

```{r import data files of Moraxella cattarrhalis and Haemophilus influenzae of gene presences absences csv from panaroo}
Morax_2_gene_preab <- 
  read.csv(file = "./panaroo_moraxella_file/Morax_gene_presence_absence_2.csv", header = TRUE, stringsAsFactors = FALSE, 
           sep = ",", 
           comment.char = "", 
           check.names = FALSE) %>% 
  tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Morax_2_gene_preab

Haemo_2_gene_preab <- 
  read.csv(file = "./panaroo_haemophilus_file/Haemo_gene_presence_absence_2.csv", header = TRUE, stringsAsFactors = FALSE, 
           sep = ",", 
           comment.char = "", 
           check.names = FALSE) %>%
  tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Haemo_2_gene_preab
```

```{r merge sample_df with Morax_2 and sample_df with Haemo_2 along with left join function}
Morax_2_gene_sample <- left_join(Morax_2_gene_preab, sample_df, by = "npg_run_id")
Morax_2_gene_sample

Haemo_2_gene_sample <- left_join(Haemo_2_gene_preab, sample_df, by = "npg_run_id")
Haemo_2_gene_sample
```

```{r calculate the number of genome count and frequency genes of Morax_2}
# Calculate number of genomes
Morax_2_gene_count <- Morax_2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Morax_2_gene_count

# Calculate Coding DNS Sequence
Morax_2_CDS_count <- Morax_2_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Morax_2_CDS_count

# Calculate frequnecy genes
Morax_2_gene_freq <- Morax_2_CDS_count %>%
  left_join(Morax_2_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
Morax_2_gene_freq
```

```{r calculate the number of genome count and frequency genes of Haemo_2}
Haemo_2_gene_count <- Haemo_2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Haemo_2_gene_count

Haemo_2_CDS_count <- Haemo_2_gene_sample %>%
  filter(CDS != "" &
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemo_2_CDS_count

Haemo_2_gene_freq <- Haemo_2_CDS_count %>%
  left_join(Haemo_2_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
Haemo_2_gene_freq
```

```{r reformat two datasets of Morax_2_gene_freq and Haemo_2_gene_freq to wide format columns}

Morax_2_gene_freq_wide <- Morax_2_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Morax_2_gene_freq_wide

Haemo_2_gene_freq_wide <- Haemo_2_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Haemo_2_gene_freq_wide
```

```{r plot scatter graph for Moraxella_2_gene in between Cambodia and Thailand after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, 
       aes(x = Thailand, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "red") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Moraxella catarrahlis 2nd\nFrequency genes by Country", 
       x = "Thailand", 
       y = "Cambodia") + 
  theme_minimal()
```

```{r plot scatter graph for Moraxella_2_gene in between United Kingdom and Cambodia after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "purple") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black") + 
  labs(title = "Moraxella catrrahlis 2nd\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph for Moraxella_2_gene in between United Kingdom and Thailand after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, 
       aes(x = Thailand, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Moraxlella catarrahlis 2nd\nFrequency genes by Country", 
       x = "Thailand", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot histogram for Moraxella_2_gene in between three conutries comparing with frequency of each genes}
ggplot(Morax_2_gene_freq, 
       aes(x = Frequency)) + 
  geom_histogram() + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Moraxella catarrahlis 2nd\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```

```{r plot scatter graph of frequency Haemophilus influenzae in between UK and Cambodia}
ggplot(Haemo_2_gene_freq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "green") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "brown") + 
  labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph of frequency Haemophilus influenzae in between UK and Thailand}
ggplot(Haemo_2_gene_freq_wide, 
       aes(x = Thailand, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "darkgrey") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "brown") + 
  labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", 
       x = "Thailand", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph of frequency Haemophilus influenzae in between Cambodia and Thailand}
ggplot(Haemo_2_gene_freq_wide, 
       aes(x = Thailand, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "darkred") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black") + 
  labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", 
       x = "Thailand", 
       y = "Cambodia") + 
  theme_minimal()
```

```{r plot histogram of Haemophilus influenzae in between three countries}
ggplot(Haemo_2_gene_freq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "blue") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Haemophilus influenzae 2nd\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```



### We will compare geneomic dataset of Moraxella cattarrhalis from South Africa, UK, Cambodia and Thailand

```{r import gene presence absence file of Moraxella cattarrhalis after running Panaroo quality control script}

Mcat_with_SA_gene_preabsence <- read.csv(
  "./Mcat_gene_presence_absence/Mcat_SA_gene_preabsence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Mcat_with_SA_gene_preabsence
```

```{r merge Mcat_SA gene presence absence with sample_df by npg_run_id }

Mcat_SA_gene_sample <- left_join(Mcat_with_SA_gene_preabsence, sample_df, by = "npg_run_id")
Mcat_SA_gene_sample

```

```{r calculate number of genomes and frequency genes for Mcat genome with dataset from South Africa, UK, Cambodia, and Thailand}
# Count distinct genomes per country
Mcat_gene_count <- Mcat_SA_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand", "South Africa")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Mcat_gene_count
# Count CDS occurrences per gene and country
Mcat_CDS_count <- Mcat_SA_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand", "South Africa")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Mcat_CDS_count

# Calculate frequency genes
Mcat_gene_freq <- Mcat_CDS_count %>%
  left_join(Mcat_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
View(Mcat_gene_freq)

```

```{r reform column of Mcat_gene_freq to wide format column}
Mcat_SA_gene_freq_wide <- Mcat_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Mcat_SA_gene_freq_wide

```

```{r plot scatter graph comparing with Cambodia and Thailand}

ggplot(Mcat_SA_gene_freq_wide, 
       aes(x = Thailand, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "green")+ 
  labs(title = "Moraxella catarrahlis with SA\nFrequency genes by Country", 
       x = "Thailand", 
       y = "Cambodia") + 
  theme_minimal()
```

```{r plot scatter graph}
ggplot(Mcat_SA_gene_freq_wide, 
       aes(x = `United Kingdom`, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "green") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "brown")+ 
  labs(title = "Moraxella catarrahlis with SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph}
ggplot(Mcat_SA_gene_freq_wide, 
       aes(x = `United Kingdom`, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "pink") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Moraxella catarrahlis with SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "Cambodia") + 
  theme_minimal()

```

```{r plot scatter graph}

ggplot(Mcat_SA_gene_freq_wide, 
       aes(x = `United Kingdom`, y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "darkgray") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "brown")+ 
  labs(title = "Moraxella catarrahlis with SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "Thailand") + 
  theme_minimal()

```

```{r plot scatter graph}

ggplot(Mcat_SA_gene_freq_wide, 
       aes(x = `South Africa`, y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "black") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "purple")+ 
  labs(title = "Moraxella catarrahlis with SA\nFrequency genes by Country", 
       x = "South Africa", 
       y = "Cambodia") + 
  theme_minimal()

```

```{r plot scatter graph }

ggplot(Mcat_SA_gene_freq_wide, 
       aes(x = `South Africa`, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "purple") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Moraxella catarrhalis with SA\nFrequency genes by Country", 
       x = "South Africa", 
       y = "Thailand") + 
  theme_minimal()

```

```{r plot histogram for Moraxella cattarrhalis in between South Africa, UK, Cambodia and Thailand}
ggplot(Mcat_gene_freq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "blue") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Mcattarrhalis with SA\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```


### We will do analysis Moraxella cattarrhalis genome after did filter comparring with only matching proportion threshold 0.95 and after running Panaroo gene presence absence csv

```{r import Mcat_version2_gene_preabsence. csv after running Panaroo}
Mcat_version2_gene_preabsence <- read.csv(
  "./Mcat_gene_presence_absence/Mcat_version2_gene_preabsence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Mcat_version2_gene_preabsence

```

```{r merge two datasets between Mcat_version2_gene_preabsence and sample_df}
Mcat_version2_gene_sample <- left_join(Mcat_version2_gene_preabsence, sample_df, by = "npg_run_id")
Mcat_version2_gene_sample
```

```{r calculate number of genomes and frequency genes for Mcattarrhalis}
# Count distinct genomes per country
Mcat_version2_genecount <- Mcat_version2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand", "South Africa")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Mcat_version2_genecount

# Count CDS occurrences per gene and country
Mcat_version2_cdscount <- Mcat_version2_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand", "South Africa")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Mcat_version2_cdscount

# Calculate frequency genes
Mcat_version2_genefreq <- Mcat_version2_cdscount %>%
  left_join(Mcat_version2_genecount, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
Mcat_version2_genefreq
```

```{r reformat Mcat_version2_genefreq to wide format columns}
Mcat_version2_genefreq_wide <- Mcat_version2_genefreq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Mcat_version2_genefreq_wide

```

```{r plot scatter graph comparing with Cambodia and Thailand}
ggplot(Mcat_version2_genefreq_wide, 
       aes(x = Thailand, 
           y = Cambodia)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "green")+ 
  labs(title = "Version2 Moraxella catarrahlis\nFrequency genes by Country", 
       x = "Thailand", 
       y = "Cambodia") + 
  theme_minimal()
```

```{r plot scatter graph comparing with Cambodia and UK}
ggplot(Mcat_version2_genefreq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "brown") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "blue")+ 
  labs(title = "Version2 Moraxella catarrahlis\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph comparing with Camboida and South Africa}
ggplot(Mcat_version2_genefreq_wide, 
       aes(x = Cambodia, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "black") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "red")+ 
  labs(title = "Version2 Moraxella catarrahlis\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph comparing with UK and South Africa}
ggplot(Mcat_version2_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "purple")+ 
  labs(title = "Version2 Moraxella catarrahlis\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph comparing with UK and Thailand}
ggplot(Mcat_version2_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "orange") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "green")+ 
  labs(title = "Version2 Moraxella catarrahlis\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot scatter graph comparing with Thailand and South Africa}
ggplot(Mcat_version2_genefreq_wide, 
       aes(x = Thailand, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Version2 Moraxella catarrahlis\nFrequency genes by Country", 
       x = "Thailand", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot histogram for Moraxella cattarrhalis comparing between UK, Cambodia, Thailand and South Africa}
ggplot(Mcat_version2_genefreq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "red") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Version2 Moraxella cattarrhalis\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))

```



### We will analize Haemophilus influenzae after filtering by N50 and Total length and running Panaroo with Haemophilus influenzae gene presence absence cvs file including South Africa data.

```{r import Haemophilus_SA_genepreabsence cvs file from Panaroo}
Haemophilus_SA_gene_preabsence <- read.csv(
  "./Haemo_SA_gene_preabsence/Haemophilus_SA_gene_preabsence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Haemophilus_SA_gene_preabsence
```

```{r merge dataset between Haemophilus_SA_gene_preabsence and sample_df}
Haemophilus_SA_gene_sample <- left_join(Haemophilus_SA_gene_preabsence, sample_df, by = "npg_run_id")
Haemophilus_SA_gene_sample

```

```{r calculate number of genomes and frequency genes for Haemophilus influenzae}
# Count distinct genomes per country
Haemophilus_SA_genecount <- Haemophilus_SA_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand", "South Africa")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Haemophilus_SA_genecount

# Count Coding DNA Sequence occurrences per gene and country
Haemophilus_SA_cdscount <- Haemophilus_SA_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand", "South Africa")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemophilus_SA_cdscount

# Calculate frequency genes
Haemophilus_SA_genefreq <- Haemophilus_SA_cdscount %>%
  left_join(Haemophilus_SA_genecount, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
Haemophilus_SA_genefreq
```

```{r reformat Haemophilus_SA_genefreq to wide format columns}
Haemophilus_SA_genefreq_wide <- Haemophilus_SA_genefreq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Haemophilus_SA_genefreq_wide

```

```{r plot scatter graph comparing between Cambodia and United Kingdom}
ggplot(Haemophilus_SA_genefreq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "green") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "gray")+ 
  labs(title = "Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph comparing between Cambodia and Thailand}
ggplot(Haemophilus_SA_genefreq_wide, 
       aes(x = Cambodia, 
           y = `Thailand`)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "darkgray")+ 
  labs(title = "Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot scatter graph comparing between Camboida and South Africa}
ggplot(Haemophilus_SA_genefreq_wide, 
       aes(x = Cambodia, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "purple") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "orange")+ 
  labs(title = "Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph comparing between United Kingdom and Thailand}
ggplot(Haemophilus_SA_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "black") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "brown")+ 
  labs(title = "Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot scatter graph comparing between United Kingdom and South Africa}
ggplot(Haemophilus_SA_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "red") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph comparing between Thailand and South Africa}
ggplot(Haemophilus_SA_genefreq_wide, 
       aes(x = Thailand, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "black") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "yellow")+ 
  labs(title = "Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "Thailand", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot histogram comparing between Cambodia, UK, Thailand and South Africa}
ggplot(Haemophilus_SA_genefreq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "lightblue") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Haemophilus influenzae including SA\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```


### We will analize Haemophilus influenzae after filtering npg_run_id by matching proportion threshold > 0.95 and then after running Panaroo with Haemophilus_version2_gene_preabsence.csv

```{r import Haemophilus version2 of gene presence absence after running Panaroo}
Haemophilus_version2_gene_preabsence <- read.csv(
  "./Haemo_SA_gene_preabsence/Haemophilus_version2_gene_preabsence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Haemophilus_version2_gene_preabsence

```

```{r merge two datasets between Haemophilus_version2_gene_preabsence and sample_df}
Haemophilus_version2_gene_sample <- left_join(Haemophilus_version2_gene_preabsence, sample_df, by = "npg_run_id")
Haemophilus_version2_gene_sample

```

```{r calculate number of genomes and frequency genes for Haemophilus influenzae version2}
# Count distinct genomes per country
Haemoinflu_version2_genecount <- Haemophilus_version2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand", "South Africa")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Haemoinflu_version2_genecount

# Count Coding DNA Sequence occurrences per gene and country
Haemoinflu_version2_cdscount <- Haemophilus_version2_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand", "South Africa")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemoinflu_version2_cdscount

# Calculate frequency genes
Haemoinflu_version2_genefreq <- Haemoinflu_version2_cdscount %>%
  left_join(Haemoinflu_version2_genecount, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
Haemoinflu_version2_genefreq
```

```{r reformat column to wide format columns}
Haemoinflu_version2_genefreq_wide <- Haemoinflu_version2_genefreq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Haemoinflu_version2_genefreq_wide
```

```{r plot scatter graph comparing between Cambodia and United Kingdom}
ggplot(Haemoinflu_version2_genefreq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "black") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "lightblue")+ 
  labs(title = "Version2 Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()

```

```{r plot scatter graph comparing between Cambodia and Thailand}
ggplot(Haemoinflu_version2_genefreq_wide, 
       aes(x = Cambodia, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "purple") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "pink")+ 
  labs(title = "Version2 Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot scatter graph comparing between United Kingdom and Thailand}
ggplot(Haemoinflu_version2_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "blue") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "red")+ 
  labs(title = "Version2 Haemophilus influenzae including SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot histogram Haemophilus influenzae version2 comparing UK, Cambodia, Thailand and South Africa}
ggplot(Haemoinflu_version2_genefreq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "blue") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Version2 Haemophilus influenzae including SA\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```

### We will analize data after filtered npg_run_id of Moraxella cattarrhalis by N50, Total length and matching proportion threshold > 0.95 and as well after running Panaroo for getting gene presence absence

```{r import 2nd Mcat version3 gene presence absence.csv file after completed running Panaroo}
Mcat_version3_gene_preabsence <- read.csv(
  "./Mcat_gene_presence_absence/2nd_Mcat_version3_gene_preabsence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Mcat_version3_gene_preabsence
```

```{r merge two datas between Mcat_version3_gene_preabsence and sample_df}
Mcat_version3_gene_sample <- left_join(Mcat_version3_gene_preabsence, sample_df, by = "npg_run_id")
Mcat_version3_gene_sample

```

```{r calculate number of genes and frequency genes of Mcattarrhalis}
# Count distinct genomes per country
Mcat_version3_genecount <- Mcat_version3_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand", "South Africa")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Mcat_version3_genecount

# Count Coding DNA Sequence occurrences per gene and country
Mcat_version3_cdscount <- Mcat_version3_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand", "South Africa")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Mcat_version3_cdscount

# Calculate frequency genes
Mcat_version3_genefreq <- Mcat_version3_cdscount %>%
  left_join(Mcat_version3_genecount, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
Mcat_version3_genefreq

```

```{r reformat columns of Mcat_version3_genefreq to wide columns format}
Mcat_version3_genefreq_wide <- Mcat_version3_genefreq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Mcat_version3_genefreq_wide

```

```{r plot scatter graph between Cambodia and Thailand}
ggplot(Mcat_version3_genefreq_wide, 
       aes(x = Cambodia, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "purple") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "black")+ 
  labs(title = "Version3 Moraxella cattarrhalis including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot scatter graph between Cambodia and United Kingdom}
ggplot(Mcat_version3_genefreq_wide, 
       aes(x = Cambodia, 
           y = `United Kingdom`)) + 
  geom_point(alpha = 0.7, 
             color = "red") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "darkgray")+ 
  labs(title = "Version3 Moraxella cattarrhalis including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "United Kingdom") + 
  theme_minimal()
```

```{r plot scatter graph between Cambodia and South Africa}
ggplot(Mcat_version3_genefreq_wide, 
       aes(x = Cambodia, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "green") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "brown")+ 
  labs(title = "Version3 Moraxella cattarrhalis including SA\nFrequency genes by Country", 
       x = "Cambodia", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph between United Kingdom and Thailand}
ggplot(Mcat_version3_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = Thailand)) + 
  geom_point(alpha = 0.7, 
             color = "brown") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "yellow")+ 
  labs(title = "Version3 Moraxella cattarrhalis including SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "Thailand") + 
  theme_minimal()
```

```{r plot scatter graph between United Kingdom and South Africa}
ggplot(Mcat_version3_genefreq_wide, 
       aes(x = `United Kingdom`, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "lightgreen") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "purple")+ 
  labs(title = "Version3 Moraxella cattarrhalis including SA\nFrequency genes by Country", 
       x = "United Kingdom", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot scatter graph between Thailand and South Africa}
ggplot(Mcat_version3_genefreq_wide, 
       aes(x = Thailand, 
           y = `South Africa`)) + 
  geom_point(alpha = 0.7, 
             color = "darkgreen") + 
  geom_abline(slope = 1, 
              intercept = 0, 
              linetype = "dashed", 
              color = "yellow")+ 
  labs(title = "Version3 Moraxella cattarrhalis including SA\nFrequency genes by Country", 
       x = "Thailand", 
       y = "South Africa") + 
  theme_minimal()
```

```{r plot histogram between United Kingdom, Cambodia, Thailand and South Africa}
ggplot(Mcat_version3_genefreq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "yellow") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Version3 Moraxella cattrrhalis including SA\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```
### We will analize Haemophilus influenzae after filtered by N50, Total length, and matching proportion threshold > 0.95 

```{r import data 2nd Hinflu version3 gene presence and absence after running Panaroo}
Hinflu_version3_gene_preabsence <- read.csv(
  "./Haemo_gene_preabsence/2nd_Hinflu_version3_gene_preabsence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
Hinflu_version3_gene_preabsence
```

```{r merge two data between Hinflu_version3_gene_preabsence and sample_df and}

```

```{r calculate number of genes and frequency genes for Hinfluenzae}

```

```{r refomat columns to wide format columns}

```

```{r plot scatter graph between Cambodia and United Kingdom}

```

```{r plot scatter graph between Cambodia and Thailand}

```

```{r plot scatter graph between United Kingdom and Thailand}

```

```{r plot histogram between United Kingdom, Cambodia and Thailand}
ggplot(Hinflu_version3_genefreq, 
       aes(x = Frequency)) + 
  geom_histogram(color = "red") + 
  facet_wrap(~ Country) + 
  theme_minimal(base_size = 14) + 
  labs(title = "Version3 Moraxella cattrrhalis including SA\nGene Frequency Distribution", 
       x = "Frequency", 
       y = "Number of Observation") + 
  theme(strip.text = element_text(face = "bold", size = 9), 
        plot.title = element_text(hjust = 0.5, face = "bold"))
```




## Compare strain frequncies between samples

## Compare gene frequencies between samples
