---
title: "NP AMR analyses"
author: "Nicholas Croucher"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(magrittr)
require(tidyverse)
require(stringr)

```

```{bash, eval=FALSE}

module load PaM/environment
module load shelf
module load bsub.py
module load irods_extractor
module load qc-short-read
iinit
export LSB_DEFAULT_USERGROUP=team284
cd /data/pam/team284/nc3/scratch/NP_AMR/data/
# QC project 6310
cd /data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc
bsub.py --queue oversubscribed --threads 1 4 6310_QC qc-short-read --studyid 6310 --kraken2_db /data/pam/software/kraken2/standard/ --read_len 151 --classification_level S -ansi-log false --kraken2_threads 16 --bracken_threads 16

rm -rf results work *_QC.[oe]; bsub.py --queue oversubscribed --threads 1 4 6310_QC qc-short-read --studyid 6310 --kraken2_db /data/pam/software/kraken2/standard/ --read_len 151 --classification_level S -ansi-log false --kraken2_threads 16 --bracken_threads 16 --type cram; watch bjobs



bsub -o output.o -e error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --input ./test_data/inputs/test_manifest.csv --read_len 150 --threshold 1 --classification_level 'G' --kraken2_threads 10 --outdir my_output

bsub.py 16 6588_QC qc-short-read --studyid 6588 --kraken2_db /data/pam/software/kraken2/standard/ --read_len 150 --classification_level S -ansi-log false

shelf get file -q import.sequencescape_study_id=6310 -q file.file_type=assembly

module load assembly-unicycler-short-read
```

Run assemblies:

```{bash, eval=FALSE}

module load PaM/environment
module load shelf
export LSB_DEFAULT_USERGROUP=team284
cd /data/pam/team284/nc3/scratch/NP_AMR/data/
```

## Load data from project

Import sample data from project

```{r Process sample data}

parse_sample_file <- function(fn) {
  
  read.table(file = fn,
             sep = "\t",
             comment.char = "",
             header = TRUE)
  
}

process_annotation_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "annotation") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",path)) %>% 
    dplyr::mutate(annotation_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_',npg_run_id)) %>% 
    dplyr::mutate(npg_run_id = gsub(".gff","",npg_run_id)) %>% 
    dplyr::select(-path)
  
}

process_assembly_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "assembly") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/velvet_assembly/contigs.fa","",gsub("/spades_assembly/contigs.fa","",path)))) %>% 
    dplyr::mutate(assembly_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_contigs.fa')) %>% 
    dplyr::select(npg_run_id,assembly_file)
  
}

process_qc_file_names <- function(df) {
  
   df %>%
    dplyr::filter(file_type == "qc") %>%
    dplyr::mutate(npg_run_id = gsub("(.)+/","",gsub("/kraken.report","",path))) %>% 
    dplyr::mutate(qc_file = paste0(stringr::str_sub(file_uuid, start = 1L, end = 6L),'_kraken.report')) %>% 
    dplyr::select(npg_run_id,qc_file)
  
}

# Parse sample files
info_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_info.tsv") %>% 
      dplyr::filter(!grepl("HiC_RMV",sample_supplier_name)) %>%
      dplyr::mutate(project = "6588"),
    parse_sample_file("./sample_data/6310_info.tsv") %>%
      dplyr::mutate(project = "6310")
  )

run_df <-
  dplyr::bind_rows(
    parse_sample_file("./sample_data/6588_runs.tsv") %>% 
      dplyr::select(-name),
    parse_sample_file("./sample_data/6310_runs.tsv") %>% 
      dplyr::select(-name)
  )

fn_6588 <- "./sample_data/6588_files.tsv"
fn_6310 <- "./sample_data/6310_files.tsv"
files_df <-
  dplyr::bind_rows(
    parse_sample_file(fn_6588) %>% process_annotation_file_names(),
    parse_sample_file(fn_6310) %>% process_annotation_file_names()
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_assembly_file_names(),
      parse_sample_file(fn_6310) %>% process_assembly_file_names()
    ),
    by = c("npg_run_id")
  ) %>% dplyr::left_join(
    dplyr::bind_rows(
      parse_sample_file(fn_6588) %>% process_qc_file_names(),
      parse_sample_file(fn_6310) %>% process_qc_file_names()
    ),
    by = c("npg_run_id")
  )

epi_df <-
  dplyr::bind_rows(
    read.csv("./sample_data/Asian_isolate_data.csv") %>% 
      dplyr::mutate(Description = gsub(" from nasopharyngeal swab","",Description)) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Species = Description) %>% 
      dplyr::mutate(Description = dplyr::if_else(Species == "Mixed bacteria","Mixed bacteria","Single isolate")),
    read.csv("./sample_data/Hampshire_isolate_data.csv") %>% 
      dplyr::mutate(Cohort = Country) %>%
      dplyr::mutate(project = "6588") %>% 
      dplyr::mutate(Description = "Single isolate"),
    read.csv("./sample_data/Global-Hinf-metadata-microreact.csv") %>% 
      dplyr::filter(disease == "") %>% 
      dplyr::select(id,country,Year) %>% 
      dplyr::filter(country == "Maela") %>% 
      dplyr::rename(Isolate = id,
                    Region = country) %>% 
      dplyr::mutate(Sample = Isolate,
                    Region = "Maela",
                    Cohort = "Maela",
                    Country = "Thailand",
                    Species = "Haemophilus influenzae",
                    project = "6310",
                    Year = as.numeric(Year)) %>% 
      dplyr::mutate(npg_run_id = gsub("(_.*?)_", "\\1#", Sample)) %>% 
      dplyr::mutate(Description = "Single isolate")
  ) %>% 
  dplyr::mutate(join_col = dplyr::if_else(project == "6588",Sample,npg_run_id)) %>% 
  dplyr::select(-c(project,npg_run_id,Sample))

# Merge data
sample_df <-
  info_df %>% 
    dplyr::left_join(run_df, by = c("sample_uuid")) %>% 
    dplyr::left_join(files_df, by = c("npg_run_id")) %>% 
    dplyr::mutate(join_col = dplyr::if_else(project == "6588",sample_donor_id,npg_run_id)) %>% 
    dplyr::left_join(epi_df, by = c("join_col")) %>% 
    dplyr::mutate(Isolate = dplyr::if_else(project == "6310",name,Isolate)) %>% 
    dplyr::select(-join_col) %>% 
    dplyr::filter(!is.na(Country))

```

## Identify species with Kraken and Bracken

Identification of species from short read data used the qc-short-read pipeline (revision 103cba2a10), which used Kraken2 (version 2.1.3) and Bracken (version 2.8). Pipeline report is [execution_report_2025-09-14_08-15-34.html](./pipeline_info/execution_report_2025-09-14_08-15-34.html).

```{bash, eval = FALSE}

module load bsub.py
module load qc-short-read
cd 6310_qc
echo -en "ID\tR1\tR2\n" > 6310.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
bsub.py -q oversubscribed 4 6310_qc_rerun qc-short-read --manifest_of_reads 6310.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false
cd ../6588_qc
echo -en "ID\tR1\tR2\n" > 6588.manifest; for f in ./results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME%_1.fastq.gz}\t${f}\t${f%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub.py -q oversubscribed 4 6588_qc_rerun qc-short-read --manifest_of_reads 6588.manifest --kraken2_db /data/pam/software/kraken2/standard/k2_standard_20250402 --read_len 150 --classification_level S --enable_building -ansi-log false

```

## Assemble reads with SPAdes and unicycler

Assembly used the assembly_unicycler_short_read pipeline (version 25.04.0), which used SPAdes (version 4.0.0) and unicycler (version 0.5.1). Pipeline reports are [execution_report_2025-09-11_20-15-21.html](./pipeline_info/execution_report_2025-09-11_20-15-21.html) and [execution_report_2025-09-11_17-22-43.html](./pipeline_info/execution_report_2025-09-11_17-22-43.html).

```{bash, eval = FALSE}

git clone --recurse-submodules git@gitlab.internal.sanger.ac.uk:sanger-pathogens/pipelines/assembly_unicycler_short_read.git
cd assembly_unicycler_short_read
echo -en "ID,R1,R2\n" > 6310.manifest; for f in ../6310_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
echo -en "ID,R1,R2\n" > 6588.manifest; for f in ../6588_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub -o 6310_output.o -e 6310_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6310.manifest --outdir 6310_assemblies
bsub -o 6588_output.o -e 6588_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6588.manifest --outdir 6588_assemblies

```

## Annotate assemblies with Bakta

The assemblies were annotated using Bakta (version 1.11).

```{bash, eval = FALSE}

module load bakta/1.11.0
declare -i N; N=1; for f in *.fa; do echo "bakta --db /data/pam/software/bakta/v6.0 --prefix ${f%.fa} --output ${f%.fa}_annotation --gram - --threads 4 ${f}" > bacta_job.${N}; chmod +x ./bacta_job.${N}; N=($N+1); done
bsub -n 4 -M 4000 -R 'select[mem>4000] rusage[mem=4000] span[hosts=1]' -o bakta.%I.o -e bakta.%I.e -J bakta_run[1-4473] ./bacta_job.\${LSB_JOBINDEX}

```

## Evaluate assemblies with Quast

Assembly statistics were calculated with Quast (version 5.0.2).

```{r Parse assembly statistics data}

# Parse assembly statistics
assembly_stats_df <-
  read.table(file = "./sample_data/NP_quast_report.tsv",
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::rename(npg_run_id = Assembly)

# Join to sample information
assembly_stats_df %>%
  dplyr::left_join(sample_df,
                   by = "npg_run_id")
```

### Merge data of two files of sample_df and assembly_stats_df
We use these two files to merge getting chronological data.

```{r, merge data, echo=TRUE}
right_df <- 
  right_join(as_tibble(assembly_stats_df),
             as_tibble(sample_df),
             by = "npg_run_id")
```

### Plot assembly data of baterial genome assemblies between total length and N50 to get quality of genome
We used total length and N50 of each genomes to find quality of genome with baseline threshold of N50 (y-axis) is 100000 bp and total length (x-axis) which is from 1.75 Mbp to 2.25 Mbp.

```{r, plot-genome assembly, include=TRUE}

ggplot(data = assembly_stats_df) + geom_point(aes(x = `Total length`, y = N50, colour = Species)) +
  geom_hline(yintercept = 100000) + scale_x_log10() + geom_vline(xintercept = c(1.75e+06, 2.25e+06))
```

### We need to plot of merging dataset to be known Quality control of each species and knowing limitted threshold to be accepted in each species.
We need to eliminate a few results from sequencing to get the better quality control of each species.

```{r, filter-data, include=TRUE}

summaried_df <- assembly_stats_df %>%
  filter(N50 > 10000 & (`Total length` > 1.5e+06 & `Total length` < 2.5e+06))

ggplot(data = summaried_df) + geom_point(aes(x = `Total length`, y = N50, colour = Species)) + geom_hline(yintercept = 80000) + geom_vline(xintercept = c(1.725e+06, 2.15e+06))
```

### Plot graph of identifing bacterial burden in each three countries

```{r, plot-graph, include=TRUE}

ggplot(data = assembly_stats_df) +
  geom_bar(aes(x = Year, fill = Country, position = "dodge")) +
  
  facet_wrap(~ Species) +
  xlab("Year of investigation") +
  ylab("Number of enrollment\nIn local hospital") +
  labs(fill = "Study Enrollment\nIn Each Countries",
       shape = "Species",
       caption = "Antimicrobial Resistance Study") +
  
  scale_shape_manual(values = c(17,19),
                     labels = c("Haemophilus influenzae", "Moraxella catarrhalis")) +
  coord_cartesian(xlim = c(2004, 2021)) +
  theme_minimal()
```

## Identify species in datasets

```{r Parse bracken outputs}

parse_bracken <- function(fn) {
  
  read.table(file = gzfile(fn),
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::filter(grepl("s__",Classification))
  
}
species_match_threshold = 10000000
species_abundance_df <-
  parse_bracken("./sample_data/6310_bracken_summary_report.tsv.gz") %>%  
  dplyr::full_join(
    parse_bracken("./sample_data/6588_bracken_summary_report.tsv.gz"),
    by = "Classification"
  ) %>%
  dplyr::mutate(Classification = gsub("(.)*\\|s__","",Classification)) %>%
  tidyr::pivot_longer(cols = -Classification, names_to = "npg_run_id", values_to = "matches") %>% 
  dplyr::rename(Species = Classification) %>% 
  dplyr::mutate(matches = dplyr::if_else(is.na(matches),0,matches)) %>% 
  dplyr::group_by(Species) %>% 
  dplyr::mutate(Total = sum(matches)) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(Total > species_match_threshold)

# Join with sample information
species_sample_df <-
  species_abundance_df %>% 
    dplyr::full_join(
      sample_df,
      by = "npg_run_id"
    )

# Plot species matches
ggplot(species_sample_df,
       aes(x = Species.x,
           y = npg_run_id,
           fill = matches+0.5)) +
  geom_tile() +
  facet_grid(Species.y ~., scales = "free_y") +
  scale_fill_continuous(trans = "log10") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

## We had filted dataset to calculate proportion and matching proportion for seeing how proportion in dataset is the right matching species

```{r create a proportion}

df <- species_sample_df %>%
  dplyr::mutate(Species.x = gsub("_"," ", Species.x)) %>%
  group_by(npg_run_id) %>%
  mutate(proportion = matches / sum(matches)) %>%
dplyr::mutate(Matching_proportion = dplyr::if_else (Species.x == Species.y, proportion, NA_real_)) %>%
tidyr::fill(Matching_proportion, .direction = "updown")
df
```

## We plot matching proportion for each types of species (Moraxella catarrhalis, Haemophilus influenzae and Mixed bacteria) to get the high quality species

```{r plot proportion}

filtered_df <- df %>%
  filter(Species.y %in%
           c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") &
           Species.y == Species.x)
summary_df <- filtered_df %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))
ggplot(filtered_df, aes(x = Species.y, y = proportion)) +
  geom_violin()+
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") +
  theme_minimal()
```

## We use the genus level to analyse Moraxella

```{r Moraxella at genus level}

genus_level_df <-
  df %>% 
    dplyr::mutate(Observed_genus = stringr::str_extract(Species.x, '[A-Za-z]+'),
                  Expected_genus = stringr::str_extract(Species.y, '[A-Za-z]+'))

filtered_genus_df <-
  genus_level_df %>%
    dplyr::filter(Species.y %in% c("Moraxella catarrhalis", "Haemophilus influenzae")) %>% 
    dplyr::group_by(npg_run_id) %>%
    dplyr::mutate(Total = sum(matches)) %>% 
    dplyr::ungroup() %>% 
    dplyr::group_by(npg_run_id,Observed_genus) %>%
    dplyr::mutate(proportion = sum(matches)/Total) %>% 
    dplyr::ungroup() %>%
    dplyr::filter(Expected_genus == Observed_genus)

ggplot(filtered_genus_df, aes(x = Expected_genus, y = proportion)) +
  geom_violin() +
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Genus", x = "Species", y = "Mean Proportion") +
  theme_minimal()
```

## We try to remove the species with lower proportion of 0.95
We want to keep samples with threshold:
N50 above 80,000, Total length above 1,725,000 bp and Total length below 2,150,000 bp, matching proportion > 0.95 of *Moraxella catarrhalis* and *Haemophilus influenzae* colony.

```{r filter the proportion }

filtered_df <- genus_level_df %>%
  filter(Species.y %in%
           c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") &
           Species.y == Species.x & proportion > 0.95)

summary_df <- filtered_df %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))

ggplot(filtered_df,
       aes(x = Species.y, y = proportion)) +
  geom_violin()+
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") +
  theme_minimal()
```

## plot the characteristic of the filter data (country and years)

```{r plot the filter dataset}

plot_genus_df <-
  filtered_genus_df %>%
  dplyr::filter(Species.y %in%
                  c("Moraxella catarrhalis", "Haemophilus influenzae") &
                Species.y == Species.x &
                proportion > 0.95) %>%
  dplyr::select(
    npg_run_id,
    Country,
    Year,
    Species.y,
    proportion,
    Matching_proportion,
    Total,
    matches) %>%
  dplyr::distinct()
plot_genus_df

ggplot(data = plot_genus_df) +
  geom_bar(aes(x = Year, fill = Country)) +
  facet_wrap(~ Species.y) +
  labs(
    title = "Total by Countries and Years",
    x = "Years",
    y = "Total",
    fill = "Country"
    ) +
  theme_minimal() +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1)
    )
```

```{r filter npg_run_id of Moraxella }

filter_Moraxella <- plot_genus_df %>%
  filter(Species.y == "Moraxella catarrhalis") %>%
select(npg_run_id) %>%
distinct()
filter_Moraxella

write.table(filter_Moraxella,
            file = "./sample_data/npg_run_id_Moraxella.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r filter npg_run_id of Haemophilus}
filter_Haemophilus <- plot_genus_df %>%
  filter(Species.y == "Haemophilus influenzae") %>%
select(npg_run_id) %>%
  distinct()
filter_Haemophilus

write.table(filter_Haemophilus,
            file = "./sample_data/npg_run_id_Haemophilus.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

## Calculate genomes of each species and each countries.

```{r calculate genomes of each species}

table(plot_genus_df$Species.y)
table(plot_genus_df$Country)
table(plot_genus_df$Year)
```

```{r Moraxella in each country and each year}

moraxella_Country_Year <-
  plot_genus_df %>%
  filter(Species.y == "Moraxella catarrhalis")

table(moraxella_Country_Year$Country,
      moraxella_Country_Year$Year)

plot_moraxella <-
  moraxella_Country_Year %>%
  count(Country, Year)

ggplot(plot_moraxella,
       aes(x = Year, y = n, fill = Country)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3, size = 3
    ) +
  ggtitle("Distribution of Moraxella catarrhalis\nin each coutries and years")
```

```{r Haemophilus influenzae in each country and each year}
Haemo_Country_Year <-
  plot_genus_df %>%
  filter(Species.y == "Haemophilus influenzae")

table(Haemo_Country_Year$Country, 
      Haemo_Country_Year$Year)
plot_Haemo <- 
  Haemo_Country_Year %>%
  count(Country, Year)
ggplot(
  plot_Haemo, aes(x = Year, y = n, fill = Country)) +
  geom_bar(stat = "identity") +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3, size = 3
    ) +
  ggtitle("Distribution of Haemophilus influenzae\nin each countries and years")
```

```{r get genomes to process}

mcat_list <-
  filtered_genus_df %>% 
    dplyr::filter(Species.x == "Moraxella catarrhalis") %>% 
    dplyr::select(npg_run_id)

write.table(mcat_list,
            file = "./sample_data/Mcat_single_isolates.txt",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r import Global Hinf metadata microreact}
Global_Hinf_metadata_microreact <- read.csv(file = "sample_data/Global-Hinf-metadata-microreact.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

```{r rename Global_Hinf_metadata_microreact dataset}
selected_Global_data <-
  Global_Hinf_metadata_microreact[, grep("^(id|Year)$", 
                                         names(Global_Hinf_metadata_microreact))]
selected_Global_data
rename_Global_data <-
  selected_Global_data %>%
  select(id, Year) %>%
  rename(
    Isolate = id, new_Year = Year
    )
rename_Global_data
```

```{r merge both datasets of Global and epi}

join_epi_Global_data <-
  left_join(
    rename_Global_data, epi_df,
            by = "Isolate"
    ) %>%
  mutate(
    Year = if_else(!is.na(new_Year),
                   new_Year, Year)
    ) %>%
  select(-new_Year)
join_epi_Global_data
```

## Use assembly statistics to QC data

```{bash, eval = FALSE}

git clone --recurse-submodules git@gitlab.internal.sanger.ac.uk:sanger-pathogens/pipelines/assembly_unicycler_short_read.git
cd assembly_unicycler_short_read
echo -en "ID,R1,R2\n" > 6310.manifest; for f in ../6310_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6310_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6310.manifest
echo -en "ID,R1,R2\n" > 6588.manifest; for f in ../6588_qc/results/fastqs/*_1.fastq.gz; do BASENAME=`basename ${f}`; echo -en "${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME},/data/pam/team284/nc3/scratch/NP_AMR/data/6588_qc/results/fastqs/${BASENAME%_1.fastq.gz}_2.fastq.gz\n"; done >> 6588.manifest
bsub -o 6310_output.o -e 6310_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6310.manifest --outdir 6310_assemblies
bsub -o 6588_output.o -e 6588_error.e -q oversubscribed -R "select[mem>4000] rusage[mem=4000]" -M4000 nextflow run . --manifest ./6588.manifest --outdir 6588_assemblies

```

## Annotate assemblies with Bakta

The assemblies were annotated using Bakta (version 1.11).

```{bash, eval = FALSE}

module load bakta/1.11.0
declare -i N; N=1; for f in *.fa; do echo "bakta --db /data/pam/software/bakta/v6.0 --prefix ${f%.fa} --output ${f%.fa}_annotation --gram - --threads 4 ${f}" > bacta_job.${N}; chmod +x ./bacta_job.${N}; N=($N+1); done
bsub -n 4 -M 4000 -R 'select[mem>4000] rusage[mem=4000] span[hosts=1]' -o bakta.%I.o -e bakta.%I.e -J bakta_run[1-4473] ./bacta_job.\${LSB_JOBINDEX}

```

## Evaluate assemblies with Quast

Assembly statistics were calculated with Quast (version 5.0.2).

```{r Parse assembly statistics data}

# Parse assembly statistics
assembly_stats_df <-
  read.table(file = "./sample_data/NP_quast_report.tsv",
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::rename(npg_run_id = Assembly)

# Join to sample information
assembly_stats_df %>%
  dplyr::left_join(sample_df,
                   by = "npg_run_id")
```

### Merge data of two files of sample_df and assembly_stats_df
We use these two files to merge getting chronological data.

```{r, merge data, echo=TRUE}
right_df <- 
  right_join(as_tibble(assembly_stats_df),
             as_tibble(sample_df),
             by = "npg_run_id")
```

### Plot assembly data of baterial genome assemblies between total length and N50 to get quality of genome
We used total length and N50 of each genomes to find quality of genome with baseline threshold of N50 (y-axis) is 100000 bp and total length (x-axis) which is from 1.75 Mbp to 2.25 Mbp.

```{r, plot-genome assembly, include=TRUE}

ggplot(data = assembly_stats_df) + geom_point(aes(x = `Total length`, y = N50, colour = Species)) +
  geom_hline(yintercept = 100000) + scale_x_log10() + geom_vline(xintercept = c(1.75e+06, 2.25e+06))
```

### We need to plot of merging dataset to be known Quality control of each species and knowing limitted threshold to be accepted in each species.
We need to eliminate a few results from sequencing to get the better quality control of each species.

```{r, filter-data, include=TRUE}

summaried_df <- assembly_stats_df %>%
  filter(N50 > 10000 & (`Total length` > 1.5e+06 & `Total length` < 2.5e+06))

ggplot(data = summaried_df) + geom_point(aes(x = `Total length`, y = N50, colour = Species)) + geom_hline(yintercept = 80000) + geom_vline(xintercept = c(1.725e+06, 2.15e+06))
```

### Plot graph of identifing bacterial burden in each three countries

```{r, plot-graph, include=TRUE}

ggplot(data = assembly_stats_df) +
  geom_bar(aes(x = Year, fill = Country, position = "dodge")) +
  
  facet_wrap(~ Species) +
  xlab("Year of investigation") +
  ylab("Number of enrollment\nIn local hospital") +
  labs(fill = "Study Enrollment\nIn Each Countries",
       shape = "Species",
       caption = "Antimicrobial Resistance Study") +
  
  scale_shape_manual(values = c(17,19),
                     labels = c("Haemophilus influenzae", "Moraxella catarrhalis")) +
  coord_cartesian(xlim = c(2004, 2021)) +
  theme_minimal()
```

## Identify species in datasets

```{r Parse bracken outputs}

parse_bracken <- function(fn) {
  
  read.table(file = gzfile(fn),
             check.names = FALSE,
             comment.char = "",
             sep ="\t",
             header = TRUE) %>% 
    dplyr::filter(grepl("s__",Classification))
  
}
species_match_threshold = 10000000
species_abundance_df <-
  parse_bracken("./sample_data/6310_bracken_summary_report.tsv.gz") %>%  
  dplyr::full_join(
    parse_bracken("./sample_data/6588_bracken_summary_report.tsv.gz"),
    by = "Classification"
  ) %>%
  dplyr::mutate(Classification = gsub("(.)*\\|s__","",Classification)) %>%
  tidyr::pivot_longer(cols = -Classification, names_to = "npg_run_id", values_to = "matches") %>% 
  dplyr::rename(Species = Classification) %>% 
  dplyr::mutate(matches = dplyr::if_else(is.na(matches),0,matches)) %>% 
  dplyr::group_by(Species) %>% 
  dplyr::mutate(Total = sum(matches)) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(Total > species_match_threshold)

# Join with sample information
species_sample_df <-
  species_abundance_df %>% 
    dplyr::full_join(
      sample_df,
      by = "npg_run_id"
    )

# Plot species matches
ggplot(species_sample_df,
       aes(x = Species.x,
           y = npg_run_id,
           fill = matches+0.5)) +
  geom_tile() +
  facet_grid(Species.y ~., scales = "free_y") +
  scale_fill_continuous(trans = "log10") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

## We had filted dataset to calculate proportion and matching proportion for seeing how proportion in dataset is the right matching species

```{r create a proportion}

df <- species_sample_df %>%
  dplyr::mutate(Species.x = gsub("_"," ", Species.x)) %>%
  group_by(npg_run_id) %>%
  mutate(proportion = matches / sum(matches)) %>%
dplyr::mutate(Matching_proportion = dplyr::if_else (Species.x == Species.y, proportion, NA_real_)) %>%
tidyr::fill(Matching_proportion, .direction = "updown")
df
```

## We plot matching proportion for each types of species (Moraxella catarrhalis, Haemophilus influenzae and Mixed bacteria) to get the high quality species

```{r plot proportion}

filtered_df <- df %>%
  filter(Species.y %in%
           c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") &
           Species.y == Species.x)
summary_df <- filtered_df %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))
ggplot(filtered_df, aes(x = Species.y, y = proportion)) +
  geom_violin()+
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") +
  theme_minimal()

```

## We use the genus level to analyse Moraxella

```{r Moraxella at genus level}

genus_level_df <-
  df %>% 
    dplyr::mutate(Observed_genus = stringr::str_extract(Species.x, '[A-Za-z]+'),
                  Expected_genus = stringr::str_extract(Species.y, '[A-Za-z]+'))

filtered_genus_df <-
  genus_level_df %>%
    dplyr::filter(Species.y %in% c("Moraxella catarrhalis", "Haemophilus influenzae")) %>% 
    dplyr::group_by(npg_run_id) %>%
    dplyr::mutate(Total = sum(matches)) %>% 
    dplyr::ungroup() %>% 
    dplyr::group_by(npg_run_id,Observed_genus) %>%
    dplyr::mutate(proportion = sum(matches)/Total) %>% 
    dplyr::ungroup() %>%
    dplyr::filter(Expected_genus == Observed_genus)

ggplot(filtered_genus_df, aes(x = Expected_genus, y = proportion)) +
  geom_violin() +
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Genus", x = "Species", y = "Mean Proportion") +
  theme_minimal()
```

## We try to remove the species with lower proportion of 0.95
We want to keep samples with threshold:
N50 above 80,000, Total length above 1,725,000 bp and Total length below 2,150,000 bp, matching proportion > 0.95 of *Moraxella catarrhalis* and *Haemophilus influenzae* colony.

```{r filter the proportion }

filtered_df <- genus_level_df %>%
  filter(Species.y %in%
           c("Moraxella catarrhalis", "Haemophilus influenzae", "Mixed bacteria") &
           Species.y == Species.x & proportion > 0.95)

summary_df <- filtered_df %>%
  group_by(Species.y) %>%
  summarise(mean_total = mean(Total, na.rm = TRUE))

ggplot(filtered_df,
       aes(x = Species.y, y = proportion)) +
  geom_violin()+
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
  labs(title = "Proportion of Each Species", x = "Species", y = "Mean Proportion") +
  theme_minimal()

```

## plot the characteristic of the filter data (country and years)

```{r plot the filter dataset}

plot_genus_df <-
  filtered_genus_df %>%
  dplyr::filter(Species.y %in%
                  c("Moraxella catarrhalis", "Haemophilus influenzae") &
                Species.y == Species.x &
                proportion > 0.95) %>%
  dplyr::select(
    npg_run_id,
    Country,
    Year,
    Species.y,
    proportion,
    Matching_proportion,
    Total,
    matches) %>%
  dplyr::distinct()
plot_genus_df

ggplot(data = plot_genus_df) +
  geom_bar(aes(x = Year, fill = Country)) +
  facet_wrap(~ Species.y) +
  labs(
    title = "Total by Countries and Years",
    x = "Years",
    y = "Total",
    fill = "Country"
    ) +
  theme_minimal() +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1)
    )

```
```{r filter npg_run_id of Moraxella }

filter_Moraxella <- plot_genus_df %>%
  filter(Species.y == "Moraxella catarrhalis") %>%
select(npg_run_id) %>%
distinct()
filter_Moraxella

write.table(filter_Moraxella,
            file = "./sample_data/npg_run_id_Moraxella.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r filter npg_run_id of Haemophilus}
filter_Haemophilus <- plot_genus_df %>%
  filter(Species.y == "Haemophilus influenzae") %>%
select(npg_run_id) %>%
  distinct()
filter_Haemophilus

write.table(filter_Haemophilus,
            file = "./sample_data/npg_run_id_Haemophilus.csv",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```
## Calculate genomes of each species and each countries.

```{r calculate genomes of each species}

table(plot_genus_df$Species.y)
table(plot_genus_df$Country)
table(plot_genus_df$Year)

```

```{r Moraxella in each country and each year}

moraxella_Country_Year <-
  plot_genus_df %>%
  filter(Species.y == "Moraxella catarrhalis")

table(moraxella_Country_Year$Country,
      moraxella_Country_Year$Year)

plot_moraxella <-
  moraxella_Country_Year %>%
  count(Country, Year)

ggplot(plot_moraxella,
       aes(x = Year, y = n, fill = Country)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3, size = 3
    ) +
  ggtitle("Distribution of Moraxella catarrhalis\nin each coutries and years")
```

```{r Haemophilus influenzae in each country and each year}
Haemo_Country_Year <-
  plot_genus_df %>%
  filter(Species.y == "Haemophilus influenzae")

table(Haemo_Country_Year$Country, 
      Haemo_Country_Year$Year)
plot_Haemo <- 
  Haemo_Country_Year %>%
  count(Country, Year)
ggplot(
  plot_Haemo, aes(x = Year, y = n, fill = Country)) +
  geom_bar(stat = "identity") +
  geom_bar(stat = "identity") +
  theme_classic() +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3, size = 3
    ) +
  ggtitle("Distribution of Haemophilus influenzae\nin each countries and years")
  
```

```{r get genomes to process}

mcat_list <-
  filtered_genus_df %>% 
    dplyr::filter(Species.x == "Moraxella catarrhalis") %>% 
    dplyr::select(npg_run_id)

write.table(mcat_list,
            file = "./sample_data/Mcat_single_isolates.txt",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE)
```

```{r import Global Hinf metadata microreact}
Global_Hinf_metadata_microreact <- read.csv(file = "sample_data/Global-Hinf-metadata-microreact.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

```{r rename Global_Hinf_metadata_microreact dataset}
selected_Global_data <-
  Global_Hinf_metadata_microreact[, grep("^(id|Year)$", 
                                         names(Global_Hinf_metadata_microreact))]
selected_Global_data
rename_Global_data <-
  selected_Global_data %>%
  select(id, Year) %>%
  rename(
    Isolate = id, new_Year = Year
    )
rename_Global_data
```

```{r merge both datasets of Global and epi}

join_epi_Global_data <-
  left_join(
    rename_Global_data, epi_df,
            by = "Isolate"
    ) %>%
  mutate(
    Year = if_else(!is.na(new_Year),
                   new_Year, Year)
    ) %>%
  select(-new_Year)
join_epi_Global_data

```

## Use AMR finder to identify resistance genes

## Use capsule typer to identify serotypes

## Use PopPUNK to categorise isolates

## Use Panaroo to identify genes
After we got the gene presence absence from Moraxella catarrhalis Panaroo, then we extracted that csv file for doing tidy column, calculate the gene frequencies and number of genomes then final plotting scatter graph and plot histograms between countries.

```{r tidy up the Moraxell gene presence absence columns to longer format columns}

Moraxella_gene_presence <- read.csv(
  "./panaroo_moraxella_file/Moraxella_gene_presence_absence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
View(Moraxella_gene_presence)
```

```{r left_join Moraxella gene presence dataset with sample_df}

Moraxella_gene_sample <- left_join(
  Moraxella_gene_presence,
  sample_df,
  by = "npg_run_id"
  )
Moraxella_gene_sample
```

```{r classifying Moraxella gene sample dataset in early and late year of UK country}

UK_early_late_df <- Moraxella_gene_sample %>%
  filter(Country == "United Kingdom",
         Year %in% c(2007:2011, 2019:2020)
         ) %>%
  mutate(Period = 
           case_when(Year >= 2007
                     & Year <= 2011 ~ "Early",
                     Year >= 2019 & Year <= 2020 ~ "Late")
         )
```

```{r calculate the number of Moraxella genome in each dataset using npg_run_id}
Genome_Count <- Moraxella_gene_sample %>%
  filter(Country %in% 
           c("United Kingdom", "Cambodia", "Thailand"),
    Species == "Moraxella catarrhalis", 
    !is.na(npg_run_id)
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id), 
            .groups = "drop")
Genome_Count
```

```{r calculate the number genome count and gene frequencies of Moraxella in dataset using npg_run_id}
Genome_Count <- Moraxella_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Genome_Count

Count_CDS <- Moraxella_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Count_CDS
gene_freq <- Count_CDS %>%
  left_join(Genome_Count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
View(gene_freq)
```

```{r reshape gene_freq to wide format}
gene_freq_wide <- gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
  pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
View(gene_freq_wide)
```

```{r plot scatter graph of frequency genes in between United Kingdom and Cambodia Country}

ggplot(gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+ labs(title = "Moraxella catarrhalis\nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Moraxella catarrhalis frequency genes in between United Kingdom and Thailand Country}
ggplot(gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") + labs(title = "Moraxella catarrhalis Frequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Moraxella frequency genes in between Cambodia and Thailand Country}
ggplot(gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+ labs(title = "Moraxella catrrhalis\nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```

```{r plot histogram of Moraxella catrrhalis frequency genes comparing with each countries}

ggplot(gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal()

```

We will calculate genome count and frequency genes of Haemophilus influenzae
```{r import Haemophilus gene presece absence csv file for tidy up the column}
Haemophilus_gene_presence <- read.csv("./panaroo_haemophilus_file/Haemo_gene_presence_absence.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, comment.char = "", check.names = FALSE) %>%
  tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Haemophilus_gene_presence
```

```{r merge Haemophilus gene presence absence and sample_df with left join function}

Haemo_gene_preab_sample <- left_join(Haemophilus_gene_presence, sample_df, by = "npg_run_id")
Haemo_gene_preab_sample

```

```{r calculate number of genome count and frequency gene of Haemophilus influenzae}
# Count distinct genomes per country
Haemo_gene_count <- Haemo_gene_preab_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Haemo_gene_count
# Count CDS occurrences per gene and country
Haemo_CDS_count <- Haemo_gene_preab_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemo_CDS_count
# Calculate frequency genes
Haemo_gene_freq <- Haemo_CDS_count %>%
  left_join(Haemo_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
View(Haemo_gene_freq)
```
```{r reform Haemo_gene_freq columns to wide format columns}
Haemo_gene_freq_wide <- Haemo_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
```

```{r plot scatter graph of Haemophilus influenzae frequency genes in between United Kingdom and Cambodia}
ggplot(Haemo_gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "red") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae \nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()

```
```{r plot scatter graph of Haemophilus influenzae frequency genes in between United Kingdom and Thailand Country}
ggplot(Haemo_gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "orange") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae \nFrequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Haemophilus influenzae frequency gene in between Cambodia and Thailand}
ggplot(Haemo_gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae \nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```
```{r plot histogram of Haemophilus influenzae with frequency genes in each countries}
ggplot(Haemo_gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal(base_size = 14) + labs(title = "Haemophilus influenzae\nGene Frequency Distribution", x = "Frequency", y = "Number of Observation") + theme(strip.text = element_text(face = "bold", size = 9), plot.title = element_text(hjust = 0.5, face = "bold"))
```
## We will extract the Moraxella catarrhalis and Haemophilus influenzae files from panaroo modifying scripts are named gene presence and absence csv to calculate the number of genomes and frequnecy genes then plot the scatter graph and histogram

```{r import Moraxella cattarrhalis of gene presences absences csv file}
Morax_2_gene_preab <- read.csv(file = "panaroo_moraxella_file/Morax_gene_presence_absence_2.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",", comment.char = "", check.names = FALSE) %>% tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Morax_2_gene_preab

Haemo_2_gene_preab <- read.csv(file = "panaroo_haemophilus_file/Haemo_gene_presence_absence_2.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",", comment.char = "", check.names = FALSE) %>%
  tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Haemo_2_gene_preab
```

```{r merge sample_df with Morax_2 and sample_df with Haemo_2 along with left join function}
Morax_2_gene_sample <- left_join(Morax_2_gene_preab, sample_df, by = "npg_run_id")
Morax_2_gene_sample

Haemo_2_gene_sample <- left_join(Haemo_2_gene_preab, sample_df, by = "npg_run_id")
Haemo_2_gene_sample
```

```{r calculate the number of genome count and frequency genes of Morax_2}
Morax_2_gene_count <- Morax_2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Morax_2_gene_count

Morax_2_CDS_count <- Morax_2_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Morax_2_CDS_count
Morax_2_gene_freq <- Morax_2_CDS_count %>%
  left_join(Morax_2_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
Morax_2_gene_freq
```
```{r calculate the number of genome count and frequency genes of Haemo_2}
Haemo_2_gene_count <- Haemo_2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Haemo_2_gene_count

Haemo_2_CDS_count <- Haemo_2_gene_sample %>%
  filter(CDS != "" &
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemo_2_CDS_count
Haemo_2_gene_freq <- Haemo_2_CDS_count %>%
  left_join(Haemo_2_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
Haemo_2_gene_freq
```

```{r reformat two datasets of Morax_2_gene_freq and Haemo_2_gene_freq to wide format columns}

Morax_2_gene_freq_wide <- Morax_2_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Morax_2_gene_freq_wide

Haemo_2_gene_freq_wide <- Haemo_2_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Haemo_2_gene_freq_wide
```
```{r plot scatter graph for Moraxella_2_gene in between Cambodia and Thailand after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "red") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Moraxella catarrahlis 2nd\nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```
```{r plot scatter graph for Moraxella_2_gene in between United Kingdom and Cambodia after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "purple") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Moraxella catrrahlis 2nd\nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph for Moraxella_2_gene in between United Kingdom and Thailand after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Moraxlella catarrahlis 2nd\nFrequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```
```{r plot histogram for Moraxella_2_gene in between three conutries comparing with frequency of each genes}
ggplot(Morax_2_gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal(base_size = 14) + labs(title = "Moraxella catarrahlis 2nd\nGene Frequency Distribution", x = "Frequency", y = "Number of Observation") + theme(strip.text = element_text(face = "bold", size = 9), plot.title = element_text(hjust = 0.5, face = "bold"))
```
```{r plot scatter graph of frequency Haemophilus influenzae in between UK and Cambodia}
ggplot(Haemo_2_gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "green") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "brown")+ labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of frequency Haemophilus influenzae in between UK and Thailand}
ggplot(Haemo_2_gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "darkgrey") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "brown")+ labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```
```{r plot scatter graph of frequency Haemophilus influenzae in between Cambodia and Thailand}
ggplot(Haemo_2_gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "darkred") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```

```{r plot histogram of Haemophilus influenzae in between three countries}
ggplot(Haemo_2_gene_freq, aes(x = Frequency)) + geom_histogram(color = "blue") + facet_wrap(~ Country) + theme_minimal(base_size = 14) + labs(title = "Haemophilus influenzae 2nd\nGene Frequency Distribution", x = "Frequency", y = "Number of Observation") + theme(strip.text = element_text(face = "bold", size = 9), plot.title = element_text(hjust = 0.5, face = "bold"))
```

After we got the gene presence absence from Moraxella catarrhalis Panaroo, then we extracted that csv file for doing tidy column, calculate the gene frequencies and number of genomes then final plotting scatter graph and plot histograms between countries.

```{r tidy up the Moraxell gene presence absence columns to longer format columns}

Moraxella_gene_presence <- read.csv(
  "./panaroo_moraxella_file/Moraxella_gene_presence_absence.csv",
  header = TRUE,
  comment.char = "",
  check.names = FALSE,
  sep = ","
  ) %>%
  tidyr::pivot_longer(
    cols = -c(
      Gene,
      `Non-unique Gene name`, 
      Annotation
      ),
    names_to = "npg_run_id",
    values_to = "CDS"
    )
View(Moraxella_gene_presence)
```

```{r left_join Moraxella gene presence dataset with sample_df}

Moraxella_gene_sample <- left_join(
  Moraxella_gene_presence,
  sample_df,
  by = "npg_run_id"
  )
Moraxella_gene_sample
```

```{r classifying Moraxella gene sample dataset in early and late year of UK country}

UK_early_late_df <- Moraxella_gene_sample %>%
  filter(Country == "United Kingdom",
         Year %in% c(2007:2011, 2019:2020)
         ) %>%
  mutate(Period = 
           case_when(Year >= 2007
                     & Year <= 2011 ~ "Early",
                     Year >= 2019 & Year <= 2020 ~ "Late")
         )
```

```{r calculate the number of Moraxella genome in each dataset using npg_run_id}
Genome_Count <- Moraxella_gene_sample %>%
  filter(Country %in% 
           c("United Kingdom", "Cambodia", "Thailand"),
    Species == "Moraxella catarrhalis", 
    !is.na(npg_run_id)
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id), 
            .groups = "drop")
Genome_Count
```

```{r calculate the number genome count and gene frequencies of Moraxella in dataset using npg_run_id}
Genome_Count <- Moraxella_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Genome_Count

Count_CDS <- Moraxella_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Count_CDS
gene_freq <- Count_CDS %>%
  left_join(Genome_Count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
View(gene_freq)
```

```{r reshape gene_freq to wide format}
gene_freq_wide <- gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
  pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
View(gene_freq_wide)
```

```{r plot scatter graph of frequency genes in between United Kingdom and Cambodia Country}

ggplot(gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+ labs(title = "Moraxella catarrhalis\nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Moraxella catarrhalis frequency genes in between United Kingdom and Thailand Country}
ggplot(gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") + labs(title = "Moraxella catarrhalis Frequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Moraxella frequency genes in between Cambodia and Thailand Country}
ggplot(gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")+ labs(title = "Moraxella catrrhalis\nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```

```{r plot histogram of Moraxella catrrhalis frequency genes comparing with each countries}
ggplot(gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal()
```

We will calculate genome count and frequency genes of Haemophilus influenzae

```{r import Haemophilus gene presece absence csv file for tidy up the column}
Haemophilus_gene_presence <- read.csv("./panaroo_haemophilus_file/Haemo_gene_presence_absence.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, comment.char = "", check.names = FALSE) %>%
  tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Haemophilus_gene_presence
```

```{r merge Haemophilus gene presence absence and sample_df with left join function}

Haemo_gene_preab_sample <- left_join(Haemophilus_gene_presence, sample_df, by = "npg_run_id")
Haemo_gene_preab_sample
```

```{r calculate number of genome count and frequency gene of Haemophilus influenzae}
# Count distinct genomes per country
Haemo_gene_count <- Haemo_gene_preab_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_count = n_distinct(npg_run_id),
    .groups = "drop")
Haemo_gene_count
# Count CDS occurrences per gene and country
Haemo_CDS_count <- Haemo_gene_preab_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemo_CDS_count
# Calculate frequency genes
Haemo_gene_freq <- Haemo_CDS_count %>%
  left_join(Haemo_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_count)
View(Haemo_gene_freq)
```

```{r reform Haemo_gene_freq columns to wide format columns}
Haemo_gene_freq_wide <- Haemo_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
```

```{r plot scatter graph of Haemophilus influenzae frequency genes in between United Kingdom and Cambodia}
ggplot(Haemo_gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "red") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae \nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Haemophilus influenzae frequency genes in between United Kingdom and Thailand Country}
ggplot(Haemo_gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "orange") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae \nFrequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of Haemophilus influenzae frequency gene in between Cambodia and Thailand}
ggplot(Haemo_gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae \nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```

```{r plot histogram of Haemophilus influenzae with frequency genes in each countries}
ggplot(Haemo_gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal(base_size = 14) + labs(title = "Haemophilus influenzae\nGene Frequency Distribution", x = "Frequency", y = "Number of Observation") + theme(strip.text = element_text(face = "bold", size = 9), plot.title = element_text(hjust = 0.5, face = "bold"))
```

## We will extract the Moraxella catarrhalis and Haemophilus influenzae files from panaroo modifying scripts are named gene presence and absence csv to calculate the number of genomes and frequnecy genes then plot the scatter graph and histogram

```{r import Moraxella cattarrhalis of gene presences absences csv file}
Morax_2_gene_preab <- read.csv(file = "panaroo_moraxella_file/Morax_gene_presence_absence_2.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",", comment.char = "", check.names = FALSE) %>% tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Morax_2_gene_preab

Haemo_2_gene_preab <- read.csv(file = "panaroo_haemophilus_file/Haemo_gene_presence_absence_2.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",", comment.char = "", check.names = FALSE) %>%
  tidyr::pivot_longer(cols = -c(Gene,`Non-unique Gene name`, Annotation), names_to = "npg_run_id", values_to = "CDS")
Haemo_2_gene_preab
```

```{r merge sample_df with Morax_2 and sample_df with Haemo_2 along with left join function}
Morax_2_gene_sample <- left_join(Morax_2_gene_preab, sample_df, by = "npg_run_id")
Morax_2_gene_sample

Haemo_2_gene_sample <- left_join(Haemo_2_gene_preab, sample_df, by = "npg_run_id")
Haemo_2_gene_sample
```

```{r calculate the number of genome count and frequency genes of Morax_2}
Morax_2_gene_count <- Morax_2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Morax_2_gene_count

Morax_2_CDS_count <- Morax_2_gene_sample %>%
  filter(CDS != "" & 
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Morax_2_CDS_count
Morax_2_gene_freq <- Morax_2_CDS_count %>%
  left_join(Morax_2_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
Morax_2_gene_freq
```
```{r calculate the number of genome count and frequency genes of Haemo_2}
Haemo_2_gene_count <- Haemo_2_gene_sample %>%
  filter(!is.na(npg_run_id),
    Country %in% c("United Kingdom", "Cambodia", "Thailand")
    ) %>%
  group_by(Country) %>%
  summarise(Genome_Count = n_distinct(npg_run_id),
    .groups = "drop")
Haemo_2_gene_count

Haemo_2_CDS_count <- Haemo_2_gene_sample %>%
  filter(CDS != "" &
         Country %in% 
           c("United Kingdom", "Cambodia", "Thailand")
         ) %>%
  group_by(Country, Gene) %>%
  summarise(CDS_count = n(), 
            .groups = "drop") %>%
arrange(Country, desc(CDS_count))
Haemo_2_CDS_count
Haemo_2_gene_freq <- Haemo_2_CDS_count %>%
  left_join(Haemo_2_gene_count, 
            by = "Country") %>%
  mutate(Frequency = CDS_count / Genome_Count)
Haemo_2_gene_freq
```

```{r reformat two datasets of Morax_2_gene_freq and Haemo_2_gene_freq to wide format columns}

Morax_2_gene_freq_wide <- Morax_2_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Morax_2_gene_freq_wide

Haemo_2_gene_freq_wide <- Haemo_2_gene_freq %>%
  select(Country, Gene, CDS_count, Genome_Count, Frequency) %>%
pivot_wider(id_cols = Gene, names_from = Country, values_from = Frequency)
Haemo_2_gene_freq_wide
```

```{r plot scatter graph for Moraxella_2_gene in between Cambodia and Thailand after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "red") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Moraxella catarrahlis 2nd\nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```

```{r plot scatter graph for Moraxella_2_gene in between United Kingdom and Cambodia after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "purple") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Moraxella catrrahlis 2nd\nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph for Moraxella_2_gene in between United Kingdom and Thailand after getting wide format columns}
ggplot(Morax_2_gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "blue") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Moraxlella catarrahlis 2nd\nFrequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```

```{r plot histogram for Moraxella_2_gene in between three conutries comparing with frequency of each genes}
ggplot(Morax_2_gene_freq, aes(x = Frequency)) + geom_histogram() + facet_wrap(~ Country) + theme_minimal(base_size = 14) + labs(title = "Moraxella catarrahlis 2nd\nGene Frequency Distribution", x = "Frequency", y = "Number of Observation") + theme(strip.text = element_text(face = "bold", size = 9), plot.title = element_text(hjust = 0.5, face = "bold"))
```

```{r plot scatter graph of frequency Haemophilus influenzae in between UK and Cambodia}
ggplot(Haemo_2_gene_freq_wide, aes(x = Cambodia, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "green") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "brown")+ labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", x = "Cambodia", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of frequency Haemophilus influenzae in between UK and Thailand}
ggplot(Haemo_2_gene_freq_wide, aes(x = Thailand, y = `United Kingdom`)) + geom_point(alpha = 0.7, color = "darkgrey") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "brown")+ labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", x = "Thailand", y = "United Kingdom") + theme_minimal()
```

```{r plot scatter graph of frequency Haemophilus influenzae in between Cambodia and Thailand}
ggplot(Haemo_2_gene_freq_wide, aes(x = Thailand, y = Cambodia)) + geom_point(alpha = 0.7, color = "darkred") + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black")+ labs(title = "Haemophilus influenzae 2nd\nFrequency genes by Country", x = "Thailand", y = "Cambodia") + theme_minimal()
```

```{r plot histogram of Haemophilus influenzae in between three countries}
ggplot(Haemo_2_gene_freq, aes(x = Frequency)) + geom_histogram(color = "blue") + facet_wrap(~ Country) + theme_minimal(base_size = 14) + labs(title = "Haemophilus influenzae 2nd\nGene Frequency Distribution", x = "Frequency", y = "Number of Observation") + theme(strip.text = element_text(face = "bold", size = 9), plot.title = element_text(hjust = 0.5, face = "bold"))
```

## Compare strain frequncies between samples

## Compare gene frequencies between samples
